### Actions

```{r, context="server"}
# Load shinyjs for enable/disable functionality
library(shinyjs)
useShinyjs()

# Source utility functions
source("source/outlier_utils.R", local = FALSE)

# Initialize reactive values at top level
prepared_changes <- reactiveVal(NULL)
```

```{r}
# Plot Data Selection
HTML('<div class="section-header">📊 Data Selection</div>')

selectizeInput("participant_16", "Participant",
  choices = participants, selected = participants[1], multiple = FALSE)
selectizeInput("trialNum_16", "Trial Number",
  choices = allTrials, selected = allTrials[4], multiple = FALSE)

HTML('<div class="section-header">📈 Plot Axes</div>')

selectizeInput("xplot_16", "X-axis",
  choices = xOptions, selected = "time", multiple = FALSE)
selectizeInput("yplot_16", "Y-axis",
  choices = xOptions, selected = "actual_pos_z", multiple = FALSE)

actionButton("refresh_plot", "🔄 Refresh Plot", class = "btn-primary")
```

```{r}
HTML('<div class="section-header">⚙️ Outlier Settings</div>')

numericInput("outlier_threshold", "Matching threshold (seconds)",
  min = 0, max = 1, value = 0.03, step = 0.001)

radioButtons("outlier_mode", "Selection Mode:", 
  choices = c("Remove Heel Strikes" = "heel","Remove Steps" = "step"),
  selected = "heel", inline = TRUE)

div(class = "helper-text",
  div("💡 Click to toggle, press TAB to switch modes")
)

checkboxInput("show_outliers_on_plot", "Show outliers on plot", value = TRUE)

HTML('<div class="section-header">📂 Import/Export</div>')

fluidRow(
  column(6,
    actionButton("load_manual_outliers", "📂 Load Outlier Files", 
                 class = "btn-info", style = "width: 100%;")
  ),
  column(6,
    actionButton("load_param_outliers", "🤖 Load Param Files", 
                 class = "btn-warning", style = "width: 100%;")
  )
)

# Include external drag & drop components
includeHTML("page16_html_imports.html")

# Custom drag & drop file inputs with enhanced UI
div(
  # Heel Strike Outliers drag & drop zone
  div(
    class = "drag-drop-zone",
    id = "heelstrike-drop-zone",
    div(class = "drag-drop-icon", "❌"),
    div(class = "drag-drop-text", strong("Import Heel Strike Outliers")),
    div(class = "drag-drop-hint", "Drag & drop CSV file here or click to browse"),
    fileInput("importHeelStrikeOutliers", NULL, 
              accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"),
              multiple = FALSE, buttonLabel = NULL, placeholder = NULL)
  ),
  
  br(),
  
  # Step Outliers drag & drop zone
  div(
    class = "drag-drop-zone",
    id = "step-drop-zone",
    div(class = "drag-drop-icon", "◐"),
    div(class = "drag-drop-text", strong("Import Step Outliers")),
    div(class = "drag-drop-hint", "Drag & drop CSV file here or click to browse"),
    fileInput("importStepOutliers", NULL,
              accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"),
              multiple = FALSE, buttonLabel = NULL, placeholder = NULL)
  )
)
```

```{r}
HTML('<div class="section-header">✅ Apply Changes</div>')

HTML('<div class="section-header">💾 Data Management</div>')

fluidRow(
  column(6, 
    actionButton("save_changes", "💾 Save Changes", class = "btn-success", 
                 style = "width: 100%;")
  ),
  column(6, 
    actionButton("save_as", "💾 Save As...", class = "btn-info", 
                 style = "width: 100%;")
  )
)

br()

HTML('<div class="section-header">🔄 Auto-save</div>')

checkboxInput("do_autosave", "Enable auto-save outliers tables", value = FALSE)
numericInput("autosave_interval", "Interval (minutes)",
             value = 5, min = 1, max = 60, step = 1)

checkboxInput("autosave_overwrite", "Overwrite default files (instead of timestamped)", value = FALSE)
checkboxInput("autosave_confirm", "Show confirmation dialog before auto-save", value = FALSE)

uiOutput("overwrite_warning")
```

Column
--------------------------------------------
### Foot Trajectories {data-width=650 data-height=800}
This interface was used to manually select the incorrectly detected steps as outliers so they could be removed from the dataset.

**Single-click** a heel strike to add it to outlier list using the selected mode above. **Double-click** to use the alternate mode without switching the radio button. It's wise to enable the auto-saving feature if you wish to reproduce this, as this takes quite some time, and the interface crashes randomly from time to time. You can also browse to the saved outliers (in the `data_extra` folder) and load them to load this list into the interface.

Also be aware that the data shown here is also based on the selection of data in the sidebar (this may already have some steps/trials filtered out).

Finally, this also includes a button on the left to save the new outliers to a `.rds` file (you can use this to overwrite the original automatic calculation of the outliers, which is far from correct).

```{r}
# Action buttons for applying changes and saving workspace
fluidRow(
  column(12,
    div(class = "action-buttons-container",
      actionButton("apply_to_workspace", "Apply Current Selection to Workspace", 
                  class = "btn-danger", 
                  style = "font-weight: bold;"),
      downloadButton("save_workspace_rds", "Save Workspace to .RDS", 
                    class = "btn-warning", 
                    style = "font-weight: bold;")
    )
  )
)

# Navigation buttons for jumping between special heel strikes and trials
fluidRow(
  column(12, 
    div(class = "navigation-buttons-container",
      # Trial navigation buttons
      actionButton("nav_prev_trial", "⬅️ Previous Trial", 
                   class = "btn-secondary", style = "margin-right: 5px;"),
      # Special step navigation buttons
      actionButton("nav_prev_special", "⬅️ Previous Special", 
                   class = "btn-info", style = "margin-right: 10px; margin-left: 5px;"),
      actionButton("nav_next_special", "➡️ Next Special", 
                   class = "btn-info", style = "margin-left: 10px; margin-right: 5px;"),
      # Trial navigation buttons
      actionButton("nav_next_trial", "➡️ Next Trial", 
                   class = "btn-secondary", style = "margin-left: 5px;"),
      br(),
      div(style = "margin-top: 8px; font-size: 12px; color: #666;",
          textOutput("nav_status", inline = TRUE))
    )
  )
)
```

```{r}
plotlyOutput("manualOutlierPlot")
```

Column
--------------------------------------------
### Heel Strike Outliers  {data-width=350 data-height=750}

**Heel Strike Removal:**
- **Yellow X markers**: Heel strikes marked for removal
- **Single-click** heel strikes when "Remove Heel Strikes" is selected, or **double-click** when "Remove Steps" is selected
- These heel strikes will be completely excluded from step calculations

```{r}
div(class = "file-info-box warning",
    div(class = "title", "Current Working File:"),
    div(class = "content", textOutput("current_hs_file", inline = TRUE))
)

DT::dataTableOutput("heelStrikeOutliersTable")
```

### Step Outliers  {data-width=350 data-height=750}

```{r}
checkboxInput("filter_outliers_table", "Filter tables by current participant/trial", value = TRUE)
```

**Step Outliers:**
- **Grey left half-circle markers**: Heel strikes marked as step outliers  
- **Single-click** heel strikes when "Remove Steps" is selected, or **double-click** when "Remove Heel Strikes" is selected
- These heel strikes will be marked as outliers in statistical analysis

```{r}
div(class = "file-info-box info",
    div(class = "title", "Current Working File:"),
    div(class = "content", textOutput("current_step_file", inline = TRUE))
)

DT::dataTableOutput("stepOutliersTable")
```


```{r, context="server"}

# Helper function to create consistent hover text for heel strikes
create_heel_strike_hover_text <- function(data, x_axis, y_axis, hs_type = "normal", icon = "", status_text = "") {
  # Safety checks
  if (is.null(data) || !is.data.frame(data) || nrow(data) == 0) return(character(0))
  
  # Check if required columns exist
  required_cols <- c("time", "participant", "trialNum", "foot", "step", x_axis, y_axis)
  missing_cols <- setdiff(required_cols, colnames(data))
  if (length(missing_cols) > 0) {
    warning(paste("Missing columns in hover text data:", paste(missing_cols, collapse = ", ")))
    return(paste("Data incomplete for", hs_type, "hover text"))
  }
  
  # Create title based on type
  title <- switch(hs_type,
    "normal" = "Heel Strike",
    "suspect" = "🟠 SUSPECT Heel Strike", 
    "heel_outlier" = "❌ HEEL STRIKE OUTLIER",
    "step_outlier" = "◐ STEP OUTLIER",
    paste(icon, "Heel Strike")
  )
  
  # Safe column access with error handling
  tryCatch({
    base_text <- paste0(
      "<b>", title, "</b><br>",
      "Time: ", round(as.numeric(data$time), 3), " s<br>",
      x_axis, ": ", round(as.numeric(data[[x_axis]]), 3), "<br>",
      y_axis, ": ", round(as.numeric(data[[y_axis]]), 3), "<br>",
      "Participant: ", as.character(data$participant), "<br>",
      "Trial: ", as.character(data$trialNum), "<br>",
      "Foot: ", as.character(data$foot), "<br>",
      "Step: ", as.character(data$step)
    )
    
    # Add status text if provided
    if (status_text != "") {
      base_text <- paste0(base_text, "<br>", "<i>Status: ", status_text, "</i>")
    }
    
    return(base_text)
  }, error = function(e) {
    warning(paste("Error creating hover text:", e$message))
    return(paste(title, "- hover text unavailable"))
  })
}

# Initialize a reactive value for plot refresh trigger
refresh_trigger <- reactiveVal(0)
last_refresh_trigger <- reactiveVal(0)

# Initialize a reactive value for tracking table content changes
table_content_hash <- reactiveVal(NULL)

# Import handlers for heel strike outliers
observeEvent(input$importHeelStrikeOutliers, {
  req(input$importHeelStrikeOutliers)
  tryCatch({
    df <- read.csv(input$importHeelStrikeOutliers$datapath, stringsAsFactors = FALSE)
    
    # Check for required columns
    required_cols <- c("participant", "trialNum", "time")
    if (all(required_cols %in% colnames(df))) {
      df$trialNum <- as.numeric(df$trialNum)
      df$participant <- as.character(df$participant)
      
      # Keep only required columns
      df <- df[, required_cols, drop = FALSE]
      
      heelstrike_outliers(df)
      last_hs_filename(basename(input$importHeelStrikeOutliers$name))
      
      # Mark as changes made since we loaded new data
      current_hash <- calculate_current_hash()
      last_saved_hs_hash(current_hash$hs)
      last_saved_step_hash(current_hash$step)
      changes_made(FALSE)  # Fresh load, no unsaved changes
      
      showNotification(paste("Heel strike outliers imported:", nrow(df), "outliers loaded from", basename(input$importHeelStrikeOutliers$name)), type = "message")
    } else {
      showNotification("Invalid CSV format. Required columns: participant, trialNum, time", type = "error")
    }
  }, error = function(e) {
    showNotification(paste("Error importing heel strike outliers:", e$message), type = "error")
  })
})

# Import handlers for step outliers  
observeEvent(input$importStepOutliers, {
  req(input$importStepOutliers)
  tryCatch({
    df <- read.csv(input$importStepOutliers$datapath, stringsAsFactors = FALSE)
    
    # Check for required columns (support both old and new format)
    if (all(c("participant", "trialNum", "time") %in% colnames(df))) {
      # New simplified format
      required_cols <- c("participant", "trialNum", "time")
      df$trialNum <- as.numeric(df$trialNum)
      df$participant <- as.character(df$participant)
      df <- df[, required_cols, drop = FALSE]
    } else if (all(c("participant", "trialNum", "time_start", "time_end") %in% colnames(df))) {
      # Old format - convert to new format using start time
      df$time <- df$time_start
      df$trialNum <- as.numeric(df$trialNum)
      df$participant <- as.character(df$participant)
      df <- df[, c("participant", "trialNum", "time"), drop = FALSE]
      showNotification("Converted old format (time_start/time_end) to new format (time)", type = "message")
    } else {
      showNotification("Invalid CSV format. Required columns: participant, trialNum, time", type = "error")
      return()
    }
    
    step_outliers(df)
    last_step_filename(basename(input$importStepOutliers$name))
    
    # Mark as fresh load, no unsaved changes
    current_hash <- calculate_current_hash()
    last_saved_hs_hash(current_hash$hs)
    last_saved_step_hash(current_hash$step)
    changes_made(FALSE)  # Fresh load, no unsaved changes
    
    showNotification(paste("Step outliers imported:", nrow(df), "outliers loaded from", basename(input$importStepOutliers$name)), type = "message")
  }, error = function(e) {
    showNotification(paste("Error importing step outliers:", e$message), type = "error")
  })
})

# Helper function to load outlier files
load_outlier_files <- function(heel_filename, step_filename, file_type = "files") {
  tryCatch({
    # Check if files exist
    heel_file <- file.path(dataExtraFolder, heel_filename)
    step_file <- file.path(dataExtraFolder, step_filename)
    
    heel_exists <- file.exists(heel_file)
    step_exists <- file.exists(step_file)
    
    if (!heel_exists && !step_exists) {
      # Detailed error message with debugging info
      debug_msg <- paste0(
        "No ", file_type, " found!\n\n",
        "Working directory: ", getwd(), "\n",
        "dataExtraFolder: ", dataExtraFolder, "\n\n",
        "Checked paths:\n",
        "• ", heel_file, " (exists: ", heel_exists, ")\n",
        "• ", step_file, " (exists: ", step_exists, ")\n\n",
        "Available files in data_extra folder:\n",
        paste(list.files(dataExtraFolder, full.names = FALSE), collapse = ", ")
      )
      showNotification(debug_msg, type = "error", duration = 15)
      return()
    }
    
    # Track loading results
    heel_loaded <- 0
    step_loaded <- 0
    errors <- c()
    
    # Load heel strike outliers if file exists
    if (heel_exists) {
      tryCatch({
        df_heel <- read.csv(heel_file, stringsAsFactors = FALSE)
        
        # Validate required columns
        if (all(c("participant", "trialNum", "time") %in% colnames(df_heel))) {
          # Clean and standardize data
          df_heel$trialNum <- as.numeric(df_heel$trialNum)
          df_heel$participant <- as.character(df_heel$participant)
          df_heel <- df_heel[, c("participant", "trialNum", "time"), drop = FALSE]
          
          # Remove any invalid rows
          df_heel <- df_heel[!is.na(df_heel$trialNum) & !is.na(df_heel$time), ]
          
          heelstrike_outliers(df_heel)
          heel_loaded <- nrow(df_heel)
          last_hs_filename(basename(heel_file))
        } else {
          errors <- c(errors, "Invalid heel strike file format")
        }
      }, error = function(e) {
        errors <- c(errors, paste("Heel strike file error:", e$message))
      })
    }
    
    # Load step outliers if file exists  
    if (step_exists) {
      tryCatch({
        df_step <- read.csv(step_file, stringsAsFactors = FALSE)
        
        # Validate required columns (support both old and new format)
        if (all(c("participant", "trialNum", "time") %in% colnames(df_step))) {
          # New simplified format
          df_step$trialNum <- as.numeric(df_step$trialNum)
          df_step$participant <- as.character(df_step$participant)
          df_step <- df_step[, c("participant", "trialNum", "time"), drop = FALSE]
        } else if (all(c("participant", "trialNum", "time_start", "time_end") %in% colnames(df_step))) {
          # Old format - convert to new format using start time
          df_step$time <- df_step$time_start
          df_step$trialNum <- as.numeric(df_step$trialNum)
          df_step$participant <- as.character(df_step$participant)
          df_step <- df_step[, c("participant", "trialNum", "time"), drop = FALSE]
        } else {
          errors <- c(errors, "Invalid step outlier file format")
          df_step <- NULL
        }
        
        if (!is.null(df_step)) {
          # Remove any invalid rows
          df_step <- df_step[!is.na(df_step$trialNum) & !is.na(df_step$time), ]
          
          step_outliers(df_step)
          step_loaded <- nrow(df_step)
          last_step_filename(basename(step_file))
        }
      }, error = function(e) {
        errors <- c(errors, paste("Step outlier file error:", e$message))
      })
    }
    
    # Show results notification
    if (length(errors) > 0) {
      showNotification(paste("Errors occurred:", paste(errors, collapse = "; ")), type = "error", duration = 10)
    } else {
      # Reset change tracking since we loaded fresh data
      current_hash <- calculate_current_hash()
      last_saved_hs_hash(current_hash$hs)
      last_saved_step_hash(current_hash$step)
      changes_made(FALSE)  # Fresh load, no unsaved changes
      
      success_msg <- paste0(
        "✅ ", tools::toTitleCase(file_type), " loaded successfully!\n",
        "❌ Heel strike outliers: ", heel_loaded, " entries\n",
        "◐ Step outliers: ", step_loaded, " entries"
      )
      showNotification(success_msg, type = "message", duration = 5)
    }
    
  }, error = function(e) {
    showNotification(paste("Error loading", file_type, ":", e$message), type = "error", duration = 10)
  })
}

# Load manual outlier files
observeEvent(input$load_manual_outliers, {
  load_outlier_files(falseHeelStrikesFile, outliersFile, "manual outlier files")
})

# Load parameter-trained outlier files
observeEvent(input$load_param_outliers, {
  load_outlier_files(falseHeelStrikesParamFile, outliersParamFile, "parameter-trained outlier files")
})

# Reactive values to store heel strike and step outliers separately
heelstrike_outliers <- reactiveVal(data.frame(
  participant = character(0),
  trialNum = numeric(0),
  time = numeric(0),
  stringsAsFactors = FALSE
))

step_outliers <- reactiveVal(data.frame(
  participant = character(0),
  trialNum = numeric(0),
  time = numeric(0),
  stringsAsFactors = FALSE
))

# Current outliers for the selected participant and trial
current_heelstrike_outliers <- reactive({
  heelstrike_outliers() %>%
    dplyr::filter(participant == input$participant_16 & trialNum == as.numeric(input$trialNum_16))
})

current_step_outliers <- reactive({
  step_outliers() %>%
    dplyr::filter(participant == input$participant_16 & trialNum == as.numeric(input$trialNum_16))
})

# Reactive values to store current axis limits (for plot memory)
axis_limits <- reactiveValues(x = NULL, y = NULL)

# --- Reactive store for last click information (to detect double clicks) ---
last_click_info <- reactiveVal(list(participant = NULL, trial = NULL, time = NA_real_, ts = 0))

# Simple plot refresh observer
observeEvent(input$refresh_plot, {
  # Show warning if there are many outliers and auto-disable if too many
  total_outliers <- 0
  if (input$show_outliers_on_plot) {
    num_hs_outliers <- nrow(heelstrike_outliers())
    num_step_outliers <- nrow(step_outliers())
    total_outliers <- num_hs_outliers + num_step_outliers
    
    if (total_outliers > 500) {
      updateCheckboxInput(session, "show_outliers_on_plot", value = FALSE)
      showNotification(paste("Auto-disabled outlier display due to", total_outliers, "outliers. Enable manually if needed."), 
                       type = "warning", duration = 5)
    } else if (total_outliers > 100) {
      showNotification(paste("Refreshing plot with", total_outliers, "outliers - this may take a moment..."), 
                       type = "message", duration = 3)
    }
  }
  
  refresh_trigger(refresh_trigger() + 1)
})

# Observe changes in the plot's layout to remember axis limits
# Only handle relayout events from the manual outlier plot
observeEvent(event_data("plotly_relayout", source = "manual_outlier_plot"), {
  relayout_data <- event_data("plotly_relayout", source = "manual_outlier_plot")
  
  if (!is.null(relayout_data)) {
    # Update x-axis limits
    if (!is.null(relayout_data[["xaxis.range[0]"]]) && !is.null(relayout_data[["xaxis.range[1]"]])) {
      axis_limits$x <- c(relayout_data[["xaxis.range[0]"]], relayout_data[["xaxis.range[1]"]])
    } else if (!is.null(relayout_data$xaxis.range)) {
      axis_limits$x <- relayout_data$xaxis.range
    }

    # Update y-axis limits
    if (!is.null(relayout_data[["yaxis.range[0]"]]) && !is.null(relayout_data[["yaxis.range[1]"]])) {
      axis_limits$y <- c(relayout_data[["yaxis.range[0]"]], relayout_data[["yaxis.range[1]"]])
    } else if (!is.null(relayout_data$yaxis.range)) {
      axis_limits$y <- relayout_data$yaxis.range
    }
  }
})


  
  # Helper function to set outlier state directly (simplified for cycling)
  handle_outlier_toggle <- function(clicked_time, current_participant, current_trial, trial_data, mode, 
                                   outlier_threshold, current_step_outliers, current_hs_outliers,
                                   step_outliers_setter, heelstrike_outliers_setter, click_type = "") {
    
    # Safety checks to ensure we have proper data frames
    if (!is.data.frame(current_step_outliers)) {
      current_step_outliers <- data.frame(
        participant = character(0), trialNum = numeric(0), time = numeric(0),
        stringsAsFactors = FALSE
      )
    }
    if (!is.data.frame(current_hs_outliers)) {
      current_hs_outliers <- data.frame(
        participant = character(0), trialNum = numeric(0), time = numeric(0),
        stringsAsFactors = FALSE
      )
    }
    
    # Add click type to notification messages
    click_suffix <- if (click_type != "") paste0(" (", click_type, ")") else ""
    
    # Find the closest heel strike
    closest_step <- find_closest_heel_strike(clicked_time, current_participant, current_trial, trial_data, outlier_threshold)
    
    if (!is.null(closest_step)) {
      # Remove from both lists first (clean slate)
      updated_hs_outliers <- current_hs_outliers %>%
        dplyr::filter(!(participant == current_participant &
                 trialNum == current_trial &
                 abs(time - closest_step$time) <= outlier_threshold))
      
      updated_step_outliers <- current_step_outliers %>%
        dplyr::filter(!(participant == current_participant &
                 trialNum == current_trial &
                 abs(time - closest_step$time) <= outlier_threshold))
      
      # Ensure results are data frames
      if (!is.data.frame(updated_hs_outliers)) {
        updated_hs_outliers <- data.frame(
          participant = character(0), trialNum = numeric(0), time = numeric(0),
          stringsAsFactors = FALSE
        )
      }
      if (!is.data.frame(updated_step_outliers)) {
        updated_step_outliers <- data.frame(
          participant = character(0), trialNum = numeric(0), time = numeric(0),
          stringsAsFactors = FALSE
        )
      }
      
      # Now set the desired state based on mode
      if (mode == "step") {
        # Add to step outliers
        new_step_outlier <- data.frame(
          participant = current_participant,
          trialNum    = current_trial,
          time        = closest_step$time,
          stringsAsFactors = FALSE
        )
        updated_step_outliers <- rbind(updated_step_outliers, new_step_outlier)
        showNotification(paste0("Marked as step outlier", click_suffix), type = "message")
      } else if (mode == "heel") {
        # Add to heel strike outliers
        new_hs_outlier <- data.frame(
          participant = current_participant,
          trialNum    = current_trial,
          time        = closest_step$time,
          stringsAsFactors = FALSE
        )
        updated_hs_outliers <- rbind(updated_hs_outliers, new_hs_outlier)
        showNotification(paste0("Marked for heel strike removal", click_suffix), type = "message")
      } else if (mode == "remove") {
        # Already removed above, just show notification
        showNotification(paste0("Removed from all outlier lists", click_suffix), type = "message")
      }
      
      # Update both lists
      heelstrike_outliers_setter(updated_hs_outliers)
      step_outliers_setter(updated_step_outliers)
    }
  }
  
  # Update outliers based on plot clicks (toggle based on selected mode)
  # Only handle clicks from the manual outlier plot, not other plotly plots
observeEvent(event_data("plotly_click", source = "manual_outlier_plot"), {
  click_data <- event_data("plotly_click", source = "manual_outlier_plot")
  
  if (!is.null(click_data) && is.list(click_data) && !is.null(click_data$x)) {
    current_participant <- input$participant_16
    current_trial <- as.numeric(input$trialNum_16)
    clicked_time <- click_data$x
    
    params <- filteredParams()
    trial_data <- params[params$participant == current_participant &
                        params$trialNum == current_trial, ]
    
    if (nrow(trial_data) == 0) return()
    
    # Find the closest heel strike to determine current state
    closest_step <- find_closest_heel_strike(clicked_time, current_participant, current_trial, trial_data, input$outlier_threshold)
    
    if (!is.null(closest_step)) {
      current_step_outliers <- step_outliers()
      current_hs_outliers <- heelstrike_outliers()
      
      # Check current state of this heel strike
      is_hs_outlier <- any(abs(current_hs_outliers$time - closest_step$time) <= input$outlier_threshold &
                          current_hs_outliers$participant == current_participant &
                          current_hs_outliers$trialNum == current_trial)
      
      is_step_outlier <- any(abs(current_step_outliers$time - closest_step$time) <= input$outlier_threshold &
                            current_step_outliers$participant == current_participant &
                            current_step_outliers$trialNum == current_trial)
      
      # Determine action per current radio selection (TAB toggles radio quickly)
      mode_to_use <- input$outlier_mode
      if ((mode_to_use == "heel" && is_hs_outlier) ||
          (mode_to_use == "step" && is_step_outlier)) {
        mode_to_use <- "remove"
        action_type <- "removed"
      } else {
        action_type <- "added"
      }
      
      # Execute the action
      handle_outlier_toggle(
        clicked_time = clicked_time,
        current_participant = current_participant,
        current_trial = current_trial,
        trial_data = trial_data,
        mode = mode_to_use,
        outlier_threshold = input$outlier_threshold,
        current_step_outliers = current_step_outliers,
        current_hs_outliers = current_hs_outliers,
        step_outliers_setter = step_outliers,
        heelstrike_outliers_setter = heelstrike_outliers,
        click_type = action_type
      )
    }
  }
})



# Navigation functionality for special heel strikes
special_heel_strikes <- reactive({
  req(filteredParams(), input$participant_16, input$trialNum_16)
  
  current_participant <- input$participant_16
  current_trial <- as.numeric(input$trialNum_16)
  
  # Get current trial data
  trial_data <- filteredParams() %>%
    dplyr::filter(participant == current_participant & trialNum == current_trial) %>%
    dplyr::arrange(time)
  
  if (nrow(trial_data) == 0) {
    return(data.frame(time = numeric(0), type = character(0), stringsAsFactors = FALSE))
  }
  
  special_times <- c()
  special_types <- c()
  
  # Add suspect steps
  if ("suspect" %in% colnames(trial_data)) {
    suspect_steps <- trial_data %>% dplyr::filter(suspect == TRUE)
    if (nrow(suspect_steps) > 0) {
      special_times <- c(special_times, suspect_steps$time)
      special_types <- c(special_types, rep("suspect", nrow(suspect_steps)))
    }
  }
  
  # Add heel strike outliers
  hs_outliers <- current_heelstrike_outliers()
  if (nrow(hs_outliers) > 0) {
    special_times <- c(special_times, hs_outliers$time)
    special_types <- c(special_types, rep("heel_strike_outlier", nrow(hs_outliers)))
  }
  
  # Add step outliers
  step_outliers_data <- current_step_outliers()
  if (nrow(step_outliers_data) > 0) {
    special_times <- c(special_times, step_outliers_data$time)
    special_types <- c(special_types, rep("step_outlier", nrow(step_outliers_data)))
  }
  
  # Create combined data frame and sort by time
  if (length(special_times) > 0) {
    special_df <- data.frame(
      time = special_times,
      type = special_types,
      stringsAsFactors = FALSE
    ) %>%
      dplyr::arrange(time) %>%
      dplyr::distinct(time, .keep_all = TRUE)  # Remove duplicates, keeping first type
  } else {
    special_df <- data.frame(time = numeric(0), type = character(0), stringsAsFactors = FALSE)
  }
  
  return(special_df)
})

# Function to navigate to a specific time using Plotly proxy (fast viewport update)
navigate_to_time <- function(target_time, current_x_axis, current_y_axis) {
  # Get current axis ranges or use data ranges if not set
  current_data <- filteredParams() %>%
    dplyr::filter(participant == input$participant_16 & trialNum == as.numeric(input$trialNum_16))
  
  if (nrow(current_data) == 0) return()
  
  # Calculate current window sizes
  if (!is.null(axis_limits$x) && length(axis_limits$x) == 2) {
    x_window_size <- axis_limits$x[2] - axis_limits$x[1]
  } else {
    # Default window size (use 10% of data range)
    x_range <- range(current_data[[current_x_axis]], na.rm = TRUE)
    x_window_size <- (x_range[2] - x_range[1]) * 0.1
  }
  
  if (!is.null(axis_limits$y) && length(axis_limits$y) == 2) {
    y_window_size <- axis_limits$y[2] - axis_limits$y[1]
  } else {
    # Default window size (use 10% of data range)
    y_range <- range(current_data[[current_y_axis]], na.rm = TRUE)
    y_window_size <- (y_range[2] - y_range[1]) * 0.1
  }
  
  # Center the windows on the target time
  x_center <- target_time
  y_center <- current_data %>%
    dplyr::filter(abs(time - target_time) == min(abs(time - target_time))) %>%
    dplyr::slice(1) %>%
    dplyr::pull(!!sym(current_y_axis))
  
  if (length(y_center) == 0) {
    y_center <- mean(current_data[[current_y_axis]], na.rm = TRUE)
  }
  
  # Calculate new axis limits
  new_x_range <- c(x_center - x_window_size/2, x_center + x_window_size/2)
  new_y_range <- c(y_center - y_window_size/2, y_center + y_window_size/2)
  
  # Update axis limits in memory for future reference
  axis_limits$x <- new_x_range
  axis_limits$y <- new_y_range
  
  # Use plotlyProxy to update just the viewport (much faster than full refresh)
  plotlyProxy("manualOutlierPlot", session) %>%
    plotlyProxyInvoke("relayout", list(
      "xaxis.range" = new_x_range,
      "yaxis.range" = new_y_range
    ))
}

# Previous special heel strike navigation
observeEvent(input$nav_prev_special, {
  req(input$xplot_16, input$yplot_16)
  
  special_data <- special_heel_strikes()
  if (nrow(special_data) == 0) {
    showNotification("No special heel strikes found for this participant/trial", type = "message")
    return()
  }
  
  # Get current center time
  current_center_time <- if (!is.null(axis_limits$x) && length(axis_limits$x) == 2) {
    mean(axis_limits$x)
  } else {
    # If no zoom set, use middle of current trial data
    trial_data <- filteredParams() %>%
      dplyr::filter(participant == input$participant_16 & trialNum == as.numeric(input$trialNum_16))
    if (nrow(trial_data) > 0) {
      mean(range(trial_data$time, na.rm = TRUE))
    } else {
      return()
    }
  }
  
  # Find previous special heel strike (at least threshold distance away)
  min_threshold <- input$outlier_threshold
  prev_specials <- special_data %>% 
    dplyr::filter(time < (current_center_time - min_threshold)) %>% 
    dplyr::arrange(desc(time))
  
  if (nrow(prev_specials) > 0) {
    target_time <- prev_specials$time[1]
    target_type <- prev_specials$type[1]
    navigate_to_time(target_time, input$xplot_16, input$yplot_16)
    showNotification(paste("Navigated to previous", gsub("_", " ", target_type), "at time", round(target_time, 3), 
                          "(≥", min_threshold, "s away)"), type = "message")
  } else {
    # Wrap around to last special heel strike
    target_time <- special_data$time[nrow(special_data)]
    target_type <- special_data$type[nrow(special_data)]
    navigate_to_time(target_time, input$xplot_16, input$yplot_16)
    showNotification(paste("Wrapped to last", gsub("_", " ", target_type), "at time", round(target_time, 3)), type = "message")
  }
})

# Next special heel strike navigation
observeEvent(input$nav_next_special, {
  req(input$xplot_16, input$yplot_16)
  
  special_data <- special_heel_strikes()
  if (nrow(special_data) == 0) {
    showNotification("No special heel strikes found for this participant/trial", type = "message")
    return()
  }
  
  # Get current center time
  current_center_time <- if (!is.null(axis_limits$x) && length(axis_limits$x) == 2) {
    mean(axis_limits$x)
  } else {
    # If no zoom set, use middle of current trial data
    trial_data <- filteredParams() %>%
      dplyr::filter(participant == input$participant_16 & trialNum == as.numeric(input$trialNum_16))
    if (nrow(trial_data) > 0) {
      mean(range(trial_data$time, na.rm = TRUE))
    } else {
      return()
    }
  }
  
  # Find next special heel strike (at least threshold distance away)
  min_threshold <- input$outlier_threshold
  next_specials <- special_data %>% 
    dplyr::filter(time > (current_center_time + min_threshold)) %>% 
    dplyr::arrange(time)
  
  if (nrow(next_specials) > 0) {
    target_time <- next_specials$time[1]
    target_type <- next_specials$type[1]
    navigate_to_time(target_time, input$xplot_16, input$yplot_16)
    showNotification(paste("Navigated to next", gsub("_", " ", target_type), "at time", round(target_time, 3), 
                          "(≥", min_threshold, "s away)"), type = "message")
  } else {
    # Wrap around to first special heel strike
    target_time <- special_data$time[1]
    target_type <- special_data$type[1]
    navigate_to_time(target_time, input$xplot_16, input$yplot_16)
    showNotification(paste("Wrapped to first", gsub("_", " ", target_type), "at time", round(target_time, 3)), type = "message")
  }
})

# Helper function to get available trials for a participant
get_participant_trials <- function(participant) {
  if (exists("allGaitParams") && !is.null(allGaitParams)) {
    participant_trials <- unique(allGaitParams$trialNum[allGaitParams$participant == participant])
    return(sort(as.numeric(participant_trials)))
  } else {
    return(sort(as.numeric(allTrials)))
  }
}

# Remember the preferred trial number across participant switches
preferred_trial <- reactiveVal(NULL)

# Update trial choices when participant changes
observeEvent(input$participant_16, {
  req(input$participant_16)
  
  # Get available trials for this participant
  available_trials <- get_participant_trials(input$participant_16)
  
  # Determine which trial to select
  selected_trial <- if (!is.null(preferred_trial()) && preferred_trial() %in% available_trials) {
    # Use the preferred trial if it's available for this participant
    preferred_trial()
  } else {
    # Fall back to first trial if preferred trial isn't available
    available_trials[1]
  }
  
  # Update the trial choices
  updateSelectizeInput(session, "trialNum_16", 
                       choices = available_trials,
                       selected = selected_trial)
}, ignoreInit = TRUE)

# Remember the preferred trial when manually changed
observeEvent(input$trialNum_16, {
  req(input$trialNum_16)
  preferred_trial(as.numeric(input$trialNum_16))
}, ignoreInit = TRUE)

# Previous trial navigation
observeEvent(input$nav_prev_trial, {
  current_participant <- input$participant_16
  current_trial <- as.numeric(input$trialNum_16)
  
  # Get available trials for current participant
  available_trials <- get_participant_trials(current_participant)
  current_trial_idx <- which(available_trials == current_trial)
  
  if (length(current_trial_idx) > 0) {
    if (current_trial_idx > 1) {
      # Go to previous trial for same participant
      new_trial <- available_trials[current_trial_idx - 1]
      updateSelectizeInput(session, "trialNum_16", selected = new_trial)
      showNotification(paste("Navigated to trial", new_trial, "for", current_participant), type = "message")
    } else {
      # Go to previous participant, last trial
      current_participant_idx <- which(participants == current_participant)
      if (length(current_participant_idx) > 0 && current_participant_idx > 1) {
        new_participant <- participants[current_participant_idx - 1]
        new_participant_trials <- get_participant_trials(new_participant)
        new_trial <- new_participant_trials[length(new_participant_trials)] # Last trial
        preferred_trial(new_trial)  # Set preferred trial before switching participant
        updateSelectizeInput(session, "participant_16", selected = new_participant)
        # Trial selection will be updated automatically by the participant observer
        showNotification(paste("Navigated to", new_participant, "trial", new_trial, "(previous participant)"), type = "message")
      } else {
        # Wrap to last participant, last trial
        new_participant <- participants[length(participants)]
        new_participant_trials <- get_participant_trials(new_participant)
        new_trial <- new_participant_trials[length(new_participant_trials)]
        preferred_trial(new_trial)  # Set preferred trial before switching participant
        updateSelectizeInput(session, "participant_16", selected = new_participant)
        # Trial selection will be updated automatically by the participant observer
        showNotification(paste("Wrapped to", new_participant, "trial", new_trial, "(last participant)"), type = "message")
      }
    }
  }
})

# Next trial navigation
observeEvent(input$nav_next_trial, {
  current_participant <- input$participant_16
  current_trial <- as.numeric(input$trialNum_16)
  
  # Get available trials for current participant
  available_trials <- get_participant_trials(current_participant)
  current_trial_idx <- which(available_trials == current_trial)
  
  if (length(current_trial_idx) > 0) {
    if (current_trial_idx < length(available_trials)) {
      # Go to next trial for same participant
      new_trial <- available_trials[current_trial_idx + 1]
      updateSelectizeInput(session, "trialNum_16", selected = new_trial)
      showNotification(paste("Navigated to trial", new_trial, "for", current_participant), type = "message")
    } else {
      # Go to next participant, first trial
      current_participant_idx <- which(participants == current_participant)
      if (length(current_participant_idx) > 0 && current_participant_idx < length(participants)) {
        new_participant <- participants[current_participant_idx + 1]
        new_participant_trials <- get_participant_trials(new_participant)
        new_trial <- new_participant_trials[1] # First trial
        preferred_trial(new_trial)  # Set preferred trial before switching participant
        updateSelectizeInput(session, "participant_16", selected = new_participant)
        # Trial selection will be updated automatically by the participant observer
        showNotification(paste("Navigated to", new_participant, "trial", new_trial, "(next participant)"), type = "message")
      } else {
        # Wrap to first participant, first trial
        new_participant <- participants[1]
        new_participant_trials <- get_participant_trials(new_participant)
        new_trial <- new_participant_trials[1]
        preferred_trial(new_trial)  # Set preferred trial before switching participant
        updateSelectizeInput(session, "participant_16", selected = new_participant)
        # Trial selection will be updated automatically by the participant observer
        showNotification(paste("Wrapped to", new_participant, "trial", new_trial, "(first participant)"), type = "message")
      }
    }
  }
})

# Status text showing current special heel strikes count
output$nav_status <- renderText({
  special_data <- special_heel_strikes()
  if (nrow(special_data) == 0) {
    "No special heel strikes found"
  } else {
    paste("Found", nrow(special_data), "special heel strikes:",
          sum(special_data$type == "suspect"), "suspect,",
          sum(special_data$type == "heel_strike_outlier"), "heel strike outliers,", 
          sum(special_data$type == "step_outlier"), "step outliers")
  }
})

# Optimize the plot rendering with caching
plot_cache <- reactiveVal(NULL)
plot_participant <- reactiveVal(NULL)
plot_trial <- reactiveVal(NULL)
plot_xaxis <- reactiveVal(NULL)
plot_yaxis <- reactiveVal(NULL)

# Render the interactive plot
output$manualOutlierPlot <- renderPlotly({
  
  # Check if we need to regenerate the base plot
  current_participant <- input$participant_16
  current_trial <- as.numeric(input$trialNum_16)
  current_xaxis <- input$xplot_16
  current_yaxis <- input$yplot_16
  
  # Only regenerate base plot if parameters changed or forced refresh
  current_refresh_trigger <- refresh_trigger()
  need_new_plot <- is.null(plot_cache()) || 
                   plot_participant() != current_participant || 
                   plot_trial() != current_trial ||
                   plot_xaxis() != current_xaxis ||
                   plot_yaxis() != current_yaxis ||
                   current_refresh_trigger != last_refresh_trigger()  # Only if refresh was triggered
  
  if (need_new_plot) {
    # Check if we have data to plot
    plot_data <- filteredParams()
    if (is.null(plot_data) || nrow(plot_data) == 0) {
      return(plot_ly() %>% 
        layout(title = "No data available for the selected participant/trial"))
    }
    
    # Create the base ggplot
    base_plot <- tryCatch({
      plot_steps(
        plot_data, current_participant, current_trial,
        current_xaxis, current_yaxis, input$filter, 
        baseSize = input$baseSize
      )
    }, error = function(e) {
      warning("Error creating base plot: ", e$message)
      return(plot_ly() %>% layout(title = paste("Error creating plot:", e$message)))
    })
    
    # Cache the base plot and parameters
    plot_cache(base_plot)
    plot_participant(current_participant)
    plot_trial(current_trial)  
    plot_xaxis(current_xaxis)
    plot_yaxis(current_yaxis)
    last_refresh_trigger(current_refresh_trigger)  # Update the last trigger value
  } else {
    # Use cached base plot
    base_plot <- plot_cache()
  }
  
  # Ensure base plot exists
  if (is.null(base_plot)) {
    return(plot_ly() %>% 
      layout(title = "Unable to generate plot"))
  }

  # Always update outlier highlighting (this is fast)
  req(filteredParams())
  
  # Get the current data for this participant and trial
  current_data <- filteredParams() %>%
    dplyr::filter(participant == current_participant & trialNum == current_trial)
  
  # Add comprehensive hover text for all heel strikes
  if (nrow(current_data) > 0) {
    current_data$plot_text <- create_heel_strike_hover_text(current_data, current_xaxis, current_yaxis, "normal")
  }
  
  # Only process outliers if they need to be shown on plot (performance optimization)
  current_data$is_hs_outlier <- FALSE
  current_hs_outliers <- NULL
  current_step_outliers_data <- NULL
  
  if (input$show_outliers_on_plot) {
    # Get current outliers for both types
    current_hs_outliers <- current_heelstrike_outliers()
    current_step_outliers_data <- current_step_outliers()
    
    # Mark heel strike outliers efficiently
    if (nrow(current_hs_outliers) > 0 && nrow(current_data) > 0) {
      threshold <- input$outlier_threshold
      
      for (i in seq_len(nrow(current_hs_outliers))) {
        within_range <- abs(current_data$time - current_hs_outliers$time[i]) <= threshold
        current_data$is_hs_outlier <- current_data$is_hs_outlier | within_range
      }
    }
  }
  
  # suspect column present
  if (!"suspect" %in% colnames(current_data)) current_data$suspect <- FALSE
  
  # Add outlier highlighting to the plot
  p <- base_plot
  
  # Add invisible overlay for comprehensive hover text on all heel strikes
  if (nrow(current_data) > 0) {
    p <- p + geom_point(data = current_data,
                       aes(x = .data[[current_xaxis]],
                           y = .data[[current_yaxis]],
                           text = plot_text),
                       alpha = 0, size = 0.1)  # Invisible but with hover text
  }
  
    # First add suspect points (orange)
  suspect_data <- current_data %>% dplyr::filter(suspect)
  if (nrow(suspect_data) > 0) {
    # Enhance hover text for suspect points
    suspect_data$plot_text <- create_heel_strike_hover_text(suspect_data, current_xaxis, current_yaxis, "suspect", "", "Flagged as suspect")
    
    p <- p + geom_point(data = suspect_data,
                       aes(x = .data[[current_xaxis]],
                           y = .data[[current_yaxis]],
                           text = plot_text),
                       color = "orange", size = 4, shape = 21, fill = "orange", alpha = 0.6)
  }
  

  
    # Add heel strike outliers (X symbol, yellow)
  hs_outlier_data <- current_data %>% dplyr::filter(is_hs_outlier)
  if (nrow(hs_outlier_data) > 0) {
    # Enhanced hover text for heel strike outliers
    hs_outlier_data$plot_text <- create_heel_strike_hover_text(hs_outlier_data, current_xaxis, current_yaxis, "heel_outlier", "", "Marked for heel strike removal")
    
    p <- p + geom_point(data = hs_outlier_data,
                         aes(x = .data[[current_xaxis]],
                             y = .data[[current_yaxis]],
                             text = plot_text),
                       color = "gold", size = 4, shape = 4, alpha = 0.8)  # Shape 4 = X, yellow color
  }
  
  # Add step outlier markers (left half circle, grey) - mark heel strikes for step outliers
  if (input$show_outliers_on_plot && !is.null(current_step_outliers_data) && nrow(current_step_outliers_data) > 0) {
    # Create markers at the heel strike times for step outliers
    step_marker_data <- current_data %>%
      dplyr::filter(
        sapply(time, function(t) {
          any(abs(current_step_outliers_data$time - t) <= input$outlier_threshold)
        })
      )
    
    if (nrow(step_marker_data) > 0) {
      # Enhanced hover text for step outliers
      step_marker_data$plot_text <- create_heel_strike_hover_text(step_marker_data, current_xaxis, current_yaxis, "step_outlier", "", "Marked as step outlier for analysis")
      
      p <- p + geom_text(data = step_marker_data,
                         aes(x = .data[[current_xaxis]],
                             y = .data[[current_yaxis]],
                             text = plot_text),
                         label = "◐", color = "grey60", size = 12, alpha = 0.6)  # Unicode left half circle, 2x larger
    }
  }
  
  # Get current axis limits (if they exist from previous zoom/pan)
  x_range_current <- isolate(axis_limits$x)
  y_range_current <- isolate(axis_limits$y)
  
  # Convert to plotly with interactive zoom and pan
  # Set source to "manual_outlier_plot" to isolate events from other plotly plots
  p <- ggplotly(p, width = input$plotwidth, height = input$plotheight, tooltip = "text", source = "manual_outlier_plot") %>%
    style(
      hoverinfo = "none",
      traces = which(sapply(.$x$data, function(d) d$type) == "scatter" & sapply(.$x$data, function(d) d$mode) == "lines")
    ) %>%
    layout(
      dragmode = "pan"
    ) %>%
    event_register("plotly_click") %>%
    event_register("plotly_relayout")
  
  # Apply remembered axis limits if they exist
  if (!is.null(x_range_current) && !is.null(y_range_current)) {
    p <- p %>%
      layout(
        xaxis = list(range = x_range_current),
        yaxis = list(range = y_range_current)
      )
  }
  
  return(p)
})



# Render the heel strike outliers table
output$heelStrikeOutliersTable <- DT::renderDataTable({
  data <- if (input$filter_outliers_table) {
    current_heelstrike_outliers()
  } else {
    heelstrike_outliers()
  }
  
  result <- create_outlier_table(data, "heelStrikeOutliersTable", "heelstrike", heelstrike_table_initialized)
  
  # Ensure we always return a table
  if (is.null(result)) {
    empty_data <- data.frame(
      Message = "No heel strike outliers found. Click on plot to add outliers.",
      stringsAsFactors = FALSE
    )
    return(DT::datatable(
      empty_data,
      options = list(dom = 't', ordering = FALSE, searching = FALSE, 
                     paging = FALSE, info = FALSE),
      selection = 'none', escape = FALSE
    ))
  }
  
  return(result)
})

# Render the step outliers table  
output$stepOutliersTable <- DT::renderDataTable({
  data <- if (input$filter_outliers_table) {
    current_step_outliers()
  } else {
    step_outliers()
  }
  
  result <- create_outlier_table(data, "stepOutliersTable", "step", step_table_initialized)
  
  # Ensure we always return a table
  if (is.null(result)) {
    empty_data <- data.frame(
      Message = "No step outliers found. Click on plot to add outliers.",
      stringsAsFactors = FALSE
    )
    return(DT::datatable(
      empty_data,
      options = list(dom = 't', ordering = FALSE, searching = FALSE, 
                     paging = FALSE, info = FALSE),
      selection = 'none', escape = FALSE
    ))
  }
  
  return(result)
})

# Track if tables have been initialized to decide between full render vs proxy update
heelstrike_table_initialized <- reactiveVal(FALSE)
step_table_initialized <- reactiveVal(FALSE)

# Reusable function to create outlier tables (optimized for performance)
create_outlier_table <- function(data, table_id, table_type, initialized_flag) {
  current_participant <- input$participant_16
  current_trial <- as.numeric(input$trialNum_16)
  
  # Track filter mode changes (only if needed)
  current_filter_mode <- input$filter_outliers_table
  filter_var_name <- paste0("last_filter_mode_", table_type)
  
  mode_changed <- FALSE
  if (!exists(filter_var_name, envir = .GlobalEnv) || 
      get(filter_var_name, envir = .GlobalEnv) != current_filter_mode) {
    assign(filter_var_name, current_filter_mode, envir = .GlobalEnv)
    initialized_flag(FALSE)
    mode_changed <- TRUE
  }
  
  # Handle empty data case efficiently
  if (is.null(data) || nrow(data) == 0) {
    empty_data <- data.frame(
      Message = paste("No", table_type, "outliers found. Click on plot to add outliers."),
      stringsAsFactors = FALSE
    )
    
    initialized_flag(TRUE)  # Mark as initialized
    return(DT::datatable(
      empty_data,
      options = list(dom = 't', ordering = FALSE, searching = FALSE, 
                     paging = FALSE, info = FALSE),
      selection = 'none', escape = FALSE
    ))
  }
  
  # Lightweight change detection - only check row count and mode changes
  simple_hash <- paste(nrow(data), current_filter_mode, current_participant, current_trial, sep = "_")
  
  hash_var_name <- paste0("table_simple_hash_", table_type)
  previous_hash <- if(exists(hash_var_name, envir = .GlobalEnv)) {
    get(hash_var_name, envir = .GlobalEnv)
  } else NULL
  
  assign(hash_var_name, simple_hash, envir = .GlobalEnv)
  needs_full_rerender <- is.null(previous_hash) || previous_hash != simple_hash
  
  # Skip re-render if no changes - but still return the existing table
  # For now, let's always render to ensure tables show up properly
  # if (initialized_flag() && !needs_full_rerender && !mode_changed) {
  #   return(NULL)
  # }
  
  # Prepare data for display
  if (table_type == "heelstrike") {
    display_data <- data %>%
      dplyr::mutate(
        Time = round(time, 4),
        unique_id = paste0("del_hs_", gsub("[^A-Za-z0-9]", "_", participant), "_", trialNum, "_", round(time * 1000)),
        Delete = paste0('<button class="btn btn-xs btn-danger" onclick="Shiny.setInputValue(\'', 
                       unique_id, '\', true, {priority: \'event\'})">Delete</button>')
      ) %>%
      dplyr::select(Participant = participant, Trial = trialNum, Time, Delete)
  } else {
    display_data <- data %>%
      dplyr::mutate(
        Time = round(time, 4),
        unique_id = paste0("del_step_", gsub("[^A-Za-z0-9]", "_", participant), "_", trialNum, "_", 
                          round(time * 1000)),
        Delete = paste0('<button class="btn btn-xs btn-danger" onclick="Shiny.setInputValue(\'', 
                       unique_id, '\', true, {priority: \'event\'})">Delete</button>')
      ) %>%
      dplyr::select(Participant = participant, Trial = trialNum, Time, Delete)
  }
  
  # Standard table background
  bg_color <- "rgba(245, 245, 245, 0.3)"
  
  # Create DataTable
  dt_table <- DT::datatable(
    display_data,
    options = list(
      pageLength = 50,
      stateSave = TRUE,
      stateDuration = 60 * 60 * 24,
      order = list(list(2, 'desc')), # Sort by time
      dom = 'frtip',
      columnDefs = list(list(targets = ncol(display_data) - 1, orderable = FALSE))
    ),
    selection = 'none',
    escape = FALSE,
    elementId = table_id
  ) %>%
    DT::formatStyle(columns = colnames(display_data), backgroundColor = bg_color)
  
  initialized_flag(TRUE)
  return(dt_table)
}

# Handle multiple point selection (updated for new outlier system)
observeEvent(input$confirm_selection, {
  selected_data <- event_data("plotly_selected", source = "manual_outlier_plot")
  
  if (!is.null(selected_data)) {
    current_participant <- input$participant_16
    current_trial <- as.numeric(input$trialNum_16)
    
    params <- filteredParams()
    trial_data <- params[params$participant == current_participant &
                        params$trialNum == current_trial, ]
    
    if (nrow(trial_data) == 0) {
      showNotification("No data found for current participant/trial", type = "warning")
      return()
    }
    
    if (input$outlier_mode == "heelstrike") {
      # Add multiple heel strike outliers
      new_hs_outliers <- lapply(seq_len(nrow(selected_data)), function(i) {
        clicked_time <- selected_data$x[i]
        
        closest_step <- find_closest_heel_strike(clicked_time, current_participant, current_trial, trial_data, input$outlier_threshold)
        
        if (!is.null(closest_step)) {
          data.frame(
            participant = closest_step$participant,
            trialNum = closest_step$trialNum,
            time = closest_step$time,
            stringsAsFactors = FALSE
          )
        } else {
          NULL
        }
      })
      
      new_hs_outliers <- do.call(rbind, new_hs_outliers)
      
      if (!is.null(new_hs_outliers) && nrow(new_hs_outliers) > 0) {
        current_hs_outliers <- heelstrike_outliers()
        updated_outliers <- rbind(current_hs_outliers, new_hs_outliers)
        updated_outliers <- unique(updated_outliers)
        heelstrike_outliers(updated_outliers)
        showNotification(paste(nrow(new_hs_outliers), "heel strike outlier(s) added"), type = "message")
      }
    } else {
      showNotification("Multi-selection for step outliers not yet implemented", type = "message")
    }
  }
})

output$save_rds <- downloadHandler(
  filename = function() {
    file.path(dataExtraFolder, paste("allGaitParams-", Sys.Date(), ".rds", sep=""))
  },
  content = function(file) {
    saveRDS(allGaitParams, file)
  }
)

# ---- Reactive values & helpers for filename-aware saving ----
last_hs_filename  <- reactiveVal(falseHeelStrikesFile)
last_step_filename <- reactiveVal(outliersFile)

# Track changes to enable/disable save button
changes_made <- reactiveVal(FALSE)
last_saved_hs_hash <- reactiveVal(NULL)
last_saved_step_hash <- reactiveVal(NULL)

# Display current working files
output$current_hs_file <- renderText({
  filename <- last_hs_filename()
  if (is.null(filename) || filename == "") {
    "No file selected"
  } else {
    filename
  }
})

output$current_step_file <- renderText({
  filename <- last_step_filename()
  if (is.null(filename) || filename == "") {
    "No file selected"
  } else {
    filename
  }
})

# Function to calculate hash of current outlier tables
calculate_current_hash <- function() {
  hs_data <- heelstrike_outliers()
  step_data <- step_outliers()
  
  # Create a simple hash based on data content
  hs_hash <- if (nrow(hs_data) > 0) {
    paste(hs_data$participant, hs_data$trialNum, hs_data$time, collapse = "_")
  } else {
    "empty"
  }
  
  step_hash <- if (nrow(step_data) > 0) {
    paste(step_data$participant, step_data$trialNum, step_data$time, collapse = "_")
  } else {
    "empty"
  }
  
  return(list(hs = hs_hash, step = step_hash))
}

# Monitor changes to outlier tables
observe({
  current_hash <- calculate_current_hash()
  
  # Check if data has changed since last save
  if (is.null(last_saved_hs_hash()) || is.null(last_saved_step_hash())) {
    # First time, assume no changes
    last_saved_hs_hash(current_hash$hs)
    last_saved_step_hash(current_hash$step)
    changes_made(FALSE)
  } else {
    # Check if current data differs from last saved
    has_changes <- (current_hash$hs != last_saved_hs_hash()) || 
                   (current_hash$step != last_saved_step_hash())
    changes_made(has_changes)
  }
})

# Update save button state based on changes
observe({
  if (changes_made()) {
    shinyjs::enable("save_changes")
    updateActionButton(session, "save_changes", 
                       label = "💾 Save Changes")
  } else {
    shinyjs::disable("save_changes")
    updateActionButton(session, "save_changes", 
                       label = "✅ No Changes")
  }
})

# Track last save moment to throttle autosave
last_save_time <- reactiveVal(Sys.time())

# Helper to save both outlier tables respecting overwrite flag
save_outlier_tables <- function(overwrite_mode = FALSE, custom_filenames = NULL){
  # Determine base filenames
  base_hs   <- last_hs_filename()  %||% falseHeelStrikesFile
  base_step <- last_step_filename() %||% outliersFile

  # Build final filenames
  if(!is.null(custom_filenames)) {
    # Use custom filenames for "Save As" functionality
    fname_hs   <- custom_filenames$hs
    fname_step <- custom_filenames$step
  } else if(overwrite_mode){
    fname_hs   <- base_hs
    fname_step <- base_step
  } else {
    stamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
    fname_hs   <- paste0(tools::file_path_sans_ext(base_hs),   "_", stamp, ".csv")
    fname_step <- paste0(tools::file_path_sans_ext(base_step), "_", stamp, ".csv")
  }

  # Get current data to save
  current_hs_data <- heelstrike_outliers()
  current_step_data <- step_outliers()
  
  # Calculate changes by comparing with existing files
  path_hs   <- file.path(dataExtraFolder, fname_hs)
  path_step <- file.path(dataExtraFolder, fname_step)
  
  # Load existing data if files exist
  existing_hs_data <- data.frame(participant = character(0), trialNum = numeric(0), time = numeric(0))
  existing_step_data <- data.frame(participant = character(0), trialNum = numeric(0), time = numeric(0))
  
  if (file.exists(path_hs)) {
    tryCatch({
      existing_hs_data <- read.csv(path_hs, stringsAsFactors = FALSE)
    }, error = function(e) {
      existing_hs_data <- data.frame(participant = character(0), trialNum = numeric(0), time = numeric(0))
    })
  }
  
  if (file.exists(path_step)) {
    tryCatch({
      existing_step_data <- read.csv(path_step, stringsAsFactors = FALSE)
    }, error = function(e) {
      existing_step_data <- data.frame(participant = character(0), trialNum = numeric(0), time = numeric(0))
    })
  }
  
  # Calculate changes for heel strike outliers
  hs_changes <- calculate_table_changes(existing_hs_data, current_hs_data)
  
  # Calculate changes for step outliers
  step_changes <- calculate_table_changes(existing_step_data, current_step_data)
  
  # Write files
  write.csv(current_hs_data, path_hs, row.names = FALSE)
  write.csv(current_step_data, path_step, row.names = FALSE)

  # Update timer
  last_save_time(Sys.time())

  list(hs_path = path_hs, step_path = path_step, hs_filename = fname_hs, step_filename = fname_step,
       hs_changes = hs_changes, step_changes = step_changes)
}

# Helper function to calculate changes between two outlier tables
calculate_table_changes <- function(old_data, new_data) {
  # Convert to comparable format (participant, trialNum, time)
  if (nrow(old_data) == 0) {
    old_key <- character(0)
  } else {
    old_key <- paste(old_data$participant, old_data$trialNum, round(old_data$time, 3), sep = "_")
  }
  
  if (nrow(new_data) == 0) {
    new_key <- character(0)
  } else {
    new_key <- paste(new_data$participant, new_data$trialNum, round(new_data$time, 3), sep = "_")
  }
  
  # Calculate added and removed
  added <- setdiff(new_key, old_key)
  removed <- setdiff(old_key, new_key)
  
  return(list(
    old_count = length(old_key),
    new_count = length(new_key),
    added = length(added),
    removed = length(removed),
    net_change = length(added) - length(removed)
  ))
}

# Common save function with detailed change reporting
perform_save_with_details <- function(overwrite_mode = TRUE, custom_filenames = NULL, 
                                     save_type = "Changes saved successfully!", 
                                     duration = 8) {
  tryCatch({
    # Save the outlier tables
    result <- save_outlier_tables(overwrite_mode = overwrite_mode, 
                                  custom_filenames = custom_filenames)
    
    # Update saved state
    current_hash <- calculate_current_hash()
    last_saved_hs_hash(current_hash$hs)
    last_saved_step_hash(current_hash$step)
    changes_made(FALSE)
    last_save_time(Sys.time())
    
    # Create detailed change report
    hs_changes <- result$hs_changes
    step_changes <- result$step_changes
    
    # Format change messages
    hs_message <- if (hs_changes$added > 0 || hs_changes$removed > 0) {
      paste0("❌ <strong>Heel Strike Outliers:</strong> ",
             if (hs_changes$added > 0) paste0("+", hs_changes$added, " added") else "",
             if (hs_changes$added > 0 && hs_changes$removed > 0) ", " else "",
             if (hs_changes$removed > 0) paste0("-", hs_changes$removed, " removed") else "",
             " (", hs_changes$old_count, " → ", hs_changes$new_count, ")")
    } else {
      paste0("❌ <strong>Heel Strike Outliers:</strong> No changes (", hs_changes$new_count, " total)")
    }
    
    step_message <- if (step_changes$added > 0 || step_changes$removed > 0) {
      paste0("◐ <strong>Step Outliers:</strong> ",
             if (step_changes$added > 0) paste0("+", step_changes$added, " added") else "",
             if (step_changes$added > 0 && step_changes$removed > 0) ", " else "",
             if (step_changes$removed > 0) paste0("-", step_changes$removed, " removed") else "",
             " (", step_changes$old_count, " → ", step_changes$new_count, ")")
    } else {
      paste0("◐ <strong>Step Outliers:</strong> No changes (", step_changes$new_count, " total)")
    }
    
    # Show success message with detailed changes
    showNotification(
      HTML(paste0(
        "✅ <strong>", save_type, "</strong><br>",
        "📁 <code>", basename(result$hs_path), "</code><br>",
        "📁 <code>", basename(result$step_path), "</code><br><br>",
        hs_message, "<br>",
        step_message
      )),
      type = "message", duration = duration
    )
    
    return(result)
    
  }, error = function(e) {
    showNotification(paste("❌ Save failed:", e$message), type = "error", duration = 8)
    return(NULL)
  })
}

# Save Changes button handler
observeEvent(input$save_changes, {
  req(changes_made())  # Only proceed if there are changes
  perform_save_with_details(overwrite_mode = TRUE, save_type = "Changes saved successfully!")
})

# Save As button handler  
observeEvent(input$save_as, {
  showModal(modalDialog(
    title = "💾 Save As - Specify Filenames",
    fluidRow(
      column(12,
        h5("Heel Strike Outliers File:"),
        textInput("save_as_hs_filename", NULL, 
                  value = paste0(tools::file_path_sans_ext(last_hs_filename() %||% falseHeelStrikesFile), "_new.csv"),
                  placeholder = "filename.csv"),
        
        h5("Step Outliers File:"),
        textInput("save_as_step_filename", NULL, 
                  value = paste0(tools::file_path_sans_ext(last_step_filename() %||% outliersFile), "_new.csv"),
                  placeholder = "filename.csv"),
        
        br(),
        div(style = "font-size: 12px; color: #6c757d;",
            "Files will be saved to: ", tags$code(dataExtraFolder))
      )
    ),
    footer = tagList(
      modalButton("❌ Cancel"),
      actionButton("confirm_save_as", "💾 Save As", class = "btn-success")
    ),
    easyClose = TRUE
  ))
})

# Confirm Save As handler
observeEvent(input$confirm_save_as, {
  req(input$save_as_hs_filename, input$save_as_step_filename)
  
  # Validate filenames
  hs_filename <- input$save_as_hs_filename
  step_filename <- input$save_as_step_filename
  
  if (!grepl("\\.csv$", hs_filename, ignore.case = TRUE)) {
    hs_filename <- paste0(hs_filename, ".csv")
  }
  if (!grepl("\\.csv$", step_filename, ignore.case = TRUE)) {
    step_filename <- paste0(step_filename, ".csv")
  }
  
  removeModal()
  
  # Update current working files to the new files
  last_hs_filename(hs_filename)
  last_step_filename(step_filename)
  
  # Use the common save function
  result <- perform_save_with_details(
    overwrite_mode = FALSE, 
    custom_filenames = list(hs = hs_filename, step = step_filename),
    save_type = paste0("Saved as new files!<br>💡 These are now your current working files")
  )
})



# --- Reactive state for auto-save ---
autosave_state <- reactiveValues(overwrite_mode = FALSE)

# Main auto-save observer (runs while auto-save is enabled)
observe({
  req(input$do_autosave)                                   # auto-save switch

  # Re-schedule this observer at the user-chosen interval
  invalidateLater(1000 * 60 * input$autosave_interval, session)

  # Throttle: don't ask again until the full interval passed since last save/skip
  if (as.numeric(difftime(Sys.time(), last_save_time(), units = "secs")) <
      input$autosave_interval * 60) return()

  # Nothing to save?  Bail out quietly
  if (!changes_made()) return()

  # Remember overwrite preference for the upcoming save
  autosave_state$overwrite_mode <- input$autosave_overwrite

  # Check if confirmation is required
  if (input$autosave_confirm) {
    # Ask the user what to do
    showModal(modalDialog(
      title = "⏱️ Auto-save triggered",
      HTML(paste0(
        "<p>The auto-save interval (", input$autosave_interval,
        " min) has elapsed.</p>",
        "<p>Do you want to save the current outlier tables now?</p>"
      )),
      footer = tagList(
        actionButton("confirm_autosave", "💾 Save now", class = "btn-success"),
        modalButton("Skip")          # closes the dialog – no save
      ),
      easyClose = TRUE
    ))
  } else {
    # Save directly without confirmation
    perform_save_with_details(
      overwrite_mode = autosave_state$overwrite_mode,
      save_type = "Auto-save completed (silent)!",
      duration = 3
    )
  }

  # Mark the inquiry time so we won't ask again until the next interval
  last_save_time(Sys.time())
})

# Handle the "Save now" button
observeEvent(input$confirm_autosave, {
  removeModal()
  # Use the common save function with detailed change reporting
  perform_save_with_details(
    overwrite_mode = autosave_state$overwrite_mode,
    save_type = "Auto-save completed!",
    duration = 5
  )
})

# Update last click information
# Only track clicks from the manual outlier plot
observe({
  click_data <- event_data("plotly_click", source = "manual_outlier_plot")
  if (!is.null(click_data)) {
    isolate({
      last_click_info(list(participant = input$participant_16,
                           trial = input$trialNum_16,
                           time = click_data$x,
                           ts = as.numeric(Sys.time())))
    })
  }
})

# Update filename trackers when files are imported or defaults loaded

# Reactive warning message for overwrite mode
output$overwrite_warning <- renderUI({
  if (input$autosave_overwrite) {
    # Show warning only when overwrite is enabled
    current_hs_file <- last_hs_filename() %||% falseHeelStrikesFile
    current_step_file <- last_step_filename() %||% outliersFile
    
    confirmation_text <- if (input$autosave_confirm) {
      "with confirmation dialog"
    } else {
      "AUTOMATICALLY without confirmation"
    }
    
    div(style = "font-size: 11px; color: #d63384; margin-top: -5px; margin-bottom: 10px; background-color: #fff3cd; padding: 5px; border-radius: 3px; border-left: 3px solid #ffc107;",
        HTML(paste0(
          "⚠️ <strong>WARNING</strong><br>",
          "Will overwrite:<br>",
          "<code>", current_hs_file, "</code><br>",
          "<code>", current_step_file, "</code><br>",
          confirmation_text
        )),
        br(),
        "✗ Unchecked: Creates timestamped backup files")
  } else {
    # Show normal info when overwrite is disabled
    confirmation_text <- if (input$autosave_confirm) {
      " (with confirmation)"
    } else {
      " (silent mode)"
    }
    
    div(style = "font-size: 11px; color: #666; margin-top: -5px; margin-bottom: 10px;",
        HTML(paste0(
          "✓ <strong>Safe mode</strong>: Creates timestamped backup files", confirmation_text
        )))
  }
})

# Observer to show confirmation dialog for applying outliers
observeEvent(input$apply_to_workspace, {
  tryCatch({
    # Ensure data.table is available
    if (!requireNamespace("data.table", quietly = TRUE)) {
      stop("Package 'data.table' is required for this operation.")
    }
    
    # Get current outliers for both types
    current_hs_outliers <- heelstrike_outliers()
    current_step_outliers <- step_outliers()
    
    if (nrow(current_hs_outliers) == 0 && nrow(current_step_outliers) == 0) {
      showNotification("No outliers to apply.", type = "message", duration = 3)
      return()
    }

    # Convert to data.table for efficient processing
    # Make copies to avoid modifying the original reactive data frames
    gait_dt <- as.data.table(allGaitParams)
    hs_outliers_dt <- as.data.table(current_hs_outliers)
    step_outliers_dt <- as.data.table(current_step_outliers)
    
    # Ensure consistent types for join columns
    gait_dt[, trialNum := as.numeric(as.character(trialNum))]
    hs_outliers_dt[, trialNum := as.numeric(as.character(trialNum))]
    step_outliers_dt[, trialNum := as.numeric(as.character(trialNum))]
    
    # --- Calculate Heel Strike Removals ---
    threshold <- input$outlier_threshold

    # Find heel strikes to remove
    hs_matches  <- find_closest_heel_strikes(current_hs_outliers, gait_dt, threshold)
    if (!is.null(hs_matches) && nrow(hs_matches) > 0) {
      hs_matches <- data.table::as.data.table(hs_matches)
      rows_to_remove <- gait_dt[hs_matches, on = .(participant, trialNum, time), nomatch = 0, which = TRUE]
    } else {
      rows_to_remove <- integer(0)
    }

    # Build summary string
    hs_impact_details <- if (length(rows_to_remove)) {
      hs_summary <- gait_dt[rows_to_remove, .N, by = .(participant, trialNum)]
      paste0(
        "❌ HEEL STRIKES TO PERMANENTLY REMOVE: ", length(rows_to_remove), " total\n   Per trial: ",
        paste0("P", hs_summary$participant, "_T", hs_summary$trialNum, " (", hs_summary$N, ")", collapse = ", "),
        "\n   ⚠️ These will be deleted from the dataset!\n\n"
      )
    } else ""
    
    # --- Calculate Step Outlier Markings ---
    step_matches <- find_closest_heel_strikes(current_step_outliers, gait_dt, threshold)
    if (!is.null(step_matches) && nrow(step_matches) > 0) {
      step_matches <- data.table::as.data.table(step_matches)
      rows_to_mark <- gait_dt[step_matches, on = .(participant, trialNum, time), nomatch = 0, which = TRUE]
    } else {
      rows_to_mark <- integer(0)
    }

    step_impact_details <- if (length(rows_to_mark)) {
      step_summary <- gait_dt[rows_to_mark, .N, by = .(participant, trialNum)]
      paste0(
        "◐ STEP OUTLIERS TO MARK: ", length(rows_to_mark), " total\n   Per trial: ",
        paste0("P", step_summary$participant, "_T", step_summary$trialNum, " (", step_summary$N, ")", collapse = ", "),
        "\n\n"
      )
    } else ""
    
    # Store the indices for the actual operation
    prepared_changes(list(remove = rows_to_remove, mark = rows_to_mark))

    # --- Create and Show Confirmation Dialog ---
    confirmation_msg <- paste0(
      "OUTLIER APPLICATION CONFIRMATION\n\n",
      hs_impact_details,
      step_impact_details,
      "⚠️  PERMANENT CHANGES TO DATASET:\n",
      "• Heel strikes will be PERMANENTLY DELETED from allGaitParams.\n",
      "• Step outliers will be marked for exclusion in analysis.\n",
      "• ALL existing outlier/suspect markings will be OVERWRITTEN.\n\n",
      "Are you sure you want to proceed?"
    )
    
    showModal(modalDialog(
      title = "⚠️ Confirm Outlier Application",
      HTML(paste0("<pre style='font-family: monospace; white-space: pre-wrap; font-size: 12px;'>", confirmation_msg, "</pre>")),
      footer = tagList(
        actionButton("confirm_apply_outliers", "✅ Apply Changes", class = "btn-success"),
        modalButton("❌ Cancel")
      ),
      easyClose = FALSE, size = "l"
    ))
    
  }, error = function(e) {
    showNotification(paste("Error preparing outlier application:", e$message), type = "error", duration = 10)
    warning("Error in apply_to_workspace preparation: ", e$message)
  })
})

# Handle confirmed outlier application
observeEvent(input$confirm_apply_outliers, {
  tryCatch({
    removeModal()

    current_hs_outliers <- heelstrike_outliers()
    current_step_outliers_data <- step_outliers()

    allGaitParams <<- apply_outliers(
      allGaitParams,
      current_hs_outliers,
      current_step_outliers_data,
      tolerance = input$outlier_threshold,
      step_tolerance = input$outlier_threshold,
      grouping_tolerance = 0.001  # Use 0.001s for grouping close steps
    )

    if (exists(".allGaitParams_dt", envir = .GlobalEnv)) {
      .allGaitParams_dt <<- NULL
    }

    refresh_trigger(refresh_trigger() + 1)

    # Show summary of what was done
    n_outlier_steps <- sum(allGaitParams$outlierSteps, na.rm = TRUE)
    n_rows_removed <- nrow(current_hs_outliers)
    
    showNotification(paste("Outliers applied successfully:", n_outlier_steps, "steps marked as outliers,", 
                         n_rows_removed, "heel strikes removed from dataset.", 
                         "All previous outlier/suspect markings have been overwritten."), 
                   type = "message", duration = 5)
  }, error = function(e) {
    showNotification(paste("Error applying outliers:", e$message), type = "error", duration = 5)
    warning("Error in confirm_apply_outliers: ", e$message)
  })
})

# Download handler for saving the entire allGaitParams object to an RDS file
output$save_workspace_rds <- downloadHandler(
  filename = function() {
    # Generate a filename with a timestamp
    paste("allGaitParams-", Sys.Date(), ".rds", sep="")
  },
  content = function(file) {
    saveRDS(allGaitParams, file)
  }
)
```

