### Actions

```{r}
tags$style(HTML("
  .section-header {
    background-color: #f8f9fa;
    padding: 8px 12px;
    margin: 10px 0 5px 0;
    border-left: 4px solid #007bff;
    font-weight: bold;
    color: #495057;
  }
  
  .btn-info {
    margin-top: 5px;
  }
"))
```

```{r}
# Plot Data Selection
HTML('<div class="section-header">📊 Data Selection</div>')

selectizeInput("participant", "Participant",
  choices = participants, selected = participants[1], multiple = FALSE)
selectizeInput("trialNum", "Trial Number",
  choices = allTrials, selected = allTrials[4], multiple = FALSE)

HTML('<div class="section-header">📈 Plot Axes</div>')

selectizeInput("xplot_16", "X-axis",
  choices = xOptions, selected = "time", multiple = FALSE)
selectizeInput("yplot_16", "Y-axis",
  choices = xOptions, selected = "pos_x", multiple = FALSE)

actionButton("refresh_plot", "🔄 Refresh Plot", class = "btn-primary")
```
```{r}
HTML('<div class="section-header">⚙️ Outlier Settings</div>')

numericInput("outlier_threshold", "Matching threshold (seconds)",
  min = 0, max = 1, value = 0.01, step = 0.001)

radioButtons("outlier_mode", "Selection Mode:", 
  choices = c("Remove Heel Strikes" = "heel","Remove Steps" = "step"),
  selected = "heel", inline = TRUE)

checkboxInput("show_outliers_on_plot", "Show outliers on plot", value = TRUE)

HTML('<div class="section-header">📂 Import/Export</div>')

actionButton("load_default_outliers", "📂 Load Default Outlier Files", 
             class = "btn-info", style = "margin-bottom: 10px; font-weight: bold;")

fileInput("importHeelStrikeOutliers", "Import Heel Strike Outliers",
  accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"))
downloadButton("export_heelstrike_outliers", "📤 Export Heel Strikes", class = "btn-secondary")

fileInput("importStepOutliers", "Import Step Outliers",
  accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"))
downloadButton("export_step_outliers", "📤 Export Steps", class = "btn-secondary")
```

```{r}
HTML('<div class="section-header">✅ Apply Changes</div>')

actionButton("overwrite_outliers", "⚡ Apply Outliers to Dataset", 
  class = "btn-success", style = "font-weight: bold;")

HTML('<div class="section-header">💾 Data Management</div>')

downloadButton("save_rds", "💾 Save Parameters to RDS", class = "btn-warning")

HTML('<div class="section-header">🔄 Auto-save</div>')

checkboxInput("do_autosave", "Enable auto-save outliers tables", value = FALSE)
numericInput("autosave_interval", "Interval (minutes)",
             value = 5, min = 1, max = 60, step = 1)

checkboxInput("autosave_overwrite", "Overwrite default files (instead of timestamped)", value = FALSE)

div(style = "font-size: 11px; color: #666; margin-top: -5px; margin-bottom: 10px;",
    "✓ Checked: Overwrites outliers_heelstrikes.csv and outliers_steps.csv",
    br(),
    "✗ Unchecked: Creates timestamped backup files")

actionButton("manual_backup", "💾 Manual Backup Both Tables", 
             class = "btn-warning", style = "margin-top: 10px; font-weight: bold;")
```


Column
--------------------------------------------
### Foot Trajectories {data-width=650 data-height=800}
This interface was used to manually select the incorrectly detected steps as outliers so they could be removed from the dataset.

Click a heel strike to add it to outlier list. It's wise to enable the auto-saving feature if you wish to reproduce this, as this takes quite some time, and the interface crashes randomly from time to time. You can also browse to the saved outliers (in the `data_extra` folder) and load them to load this list into the interface.

Also be aware that the data shown here is also based on the selection of data in the sidebar (this may already have some steps/trials filtered out).

Finally, this also includes a button on the left to save the new outliers to a `.rds` file (you can use this to overwrite the original automatic calculation of the outliers, which is far from correct).

```{r}
# Navigation buttons for jumping between special heel strikes
fluidRow(
  column(12, 
    div(style = "text-align: center; margin: 15px 0; padding: 10px; background-color: #f8f9fa; border-radius: 5px;",
      actionButton("nav_prev_special", "⬅️ Previous Special", 
                   class = "btn-info", style = "margin-right: 10px;"),
      actionButton("nav_next_special", "➡️ Next Special", 
                   class = "btn-info", style = "margin-left: 10px;"),
      br(),
      div(style = "margin-top: 8px; font-size: 12px; color: #666;",
          textOutput("nav_status", inline = TRUE))
    )
  )
)
```

```{r}
plotlyOutput("manualOutlierPlot")
```

Column
--------------------------------------------
### Heel Strike Outliers  {data-width=350 data-height=750}

**Heel Strike Removal:**
- **Yellow X markers**: Heel strikes marked for removal
- Click heel strikes on the plot to add/remove from this list  
- These heel strikes will be completely excluded from step calculations

```{r}
DT::dataTableOutput("heelStrikeOutliersTable")
```

### Step Outliers  {data-width=350 data-height=750}

```{r}
checkboxInput("filter_outliers_table", "Filter tables by current participant/trial", value = TRUE)
```

**Step Outliers:**
- **Grey left half-circle markers**: Heel strikes marked as step outliers  
- Click heel strikes on the plot to add/remove from this list
- These heel strikes will be marked as outliers in statistical analysis

```{r}
DT::dataTableOutput("stepOutliersTable")
```


```{r, context="server"}

# Helper function to create consistent hover text for heel strikes
create_heel_strike_hover_text <- function(data, x_axis, y_axis, hs_type = "normal", icon = "", status_text = "") {
  if (nrow(data) == 0) return(character(0))
  
  # Create title based on type
  title <- switch(hs_type,
    "normal" = "Heel Strike",
    "suspect" = "🟠 SUSPECT Heel Strike", 
    "heel_outlier" = "❌ HEEL STRIKE OUTLIER",
    "step_outlier" = "◐ STEP OUTLIER",
    paste(icon, "Heel Strike")
  )
  
  base_text <- paste0(
    "<b>", title, "</b><br>",
    "Time: ", round(data$time, 3), " s<br>",
    x_axis, ": ", round(data[[x_axis]], 3), "<br>",
    y_axis, ": ", round(data[[y_axis]], 3), "<br>",
    "Participant: ", data$participant, "<br>",
    "Trial: ", data$trialNum, "<br>",
    "Foot: ", data$foot, "<br>",
    "Step: ", data$step
  )
  
  # Add status text if provided
  if (status_text != "") {
    base_text <- paste0(base_text, "<br>", "<i>Status: ", status_text, "</i>")
  }
  
  return(base_text)
}

# Initialize a reactive value for plot refresh trigger
refresh_trigger <- reactiveVal(0)
last_refresh_trigger <- reactiveVal(0)

# Initialize a reactive value for tracking table content changes
table_content_hash <- reactiveVal(NULL)

# Import handlers for heel strike outliers
observeEvent(input$importHeelStrikeOutliers, {
  req(input$importHeelStrikeOutliers)
  tryCatch({
    df <- read.csv(input$importHeelStrikeOutliers$datapath, stringsAsFactors = FALSE)
    
    # Check for required columns
    required_cols <- c("participant", "trialNum", "time")
    if (all(required_cols %in% colnames(df))) {
      df$trialNum <- as.numeric(df$trialNum)
      df$participant <- as.character(df$participant)
      
      # Keep only required columns
      df <- df[, required_cols, drop = FALSE]
      
      heelstrike_outliers(df)
      showNotification(paste("Heel strike outliers imported:", nrow(df), "outliers loaded"), type = "message")
    } else {
      showNotification("Invalid CSV format. Required columns: participant, trialNum, time", type = "error")
    }
  }, error = function(e) {
    showNotification(paste("Error importing heel strike outliers:", e$message), type = "error")
  })
})

# Import handlers for step outliers  
observeEvent(input$importStepOutliers, {
  req(input$importStepOutliers)
  tryCatch({
    df <- read.csv(input$importStepOutliers$datapath, stringsAsFactors = FALSE)
    
    # Check for required columns (support both old and new format)
    if (all(c("participant", "trialNum", "time") %in% colnames(df))) {
      # New simplified format
      required_cols <- c("participant", "trialNum", "time")
      df$trialNum <- as.numeric(df$trialNum)
      df$participant <- as.character(df$participant)
      df <- df[, required_cols, drop = FALSE]
    } else if (all(c("participant", "trialNum", "time_start", "time_end") %in% colnames(df))) {
      # Old format - convert to new format using start time
      df$time <- df$time_start
      df$trialNum <- as.numeric(df$trialNum)
      df$participant <- as.character(df$participant)
      df <- df[, c("participant", "trialNum", "time"), drop = FALSE]
      showNotification("Converted old format (time_start/time_end) to new format (time)", type = "message")
    } else {
      showNotification("Invalid CSV format. Required columns: participant, trialNum, time", type = "error")
      return()
    }
    
    step_outliers(df)
    showNotification(paste("Step outliers imported:", nrow(df), "outliers loaded"), type = "message")
  }, error = function(e) {
    showNotification(paste("Error importing step outliers:", e$message), type = "error")
  })
})

# Load default outlier files
observeEvent(input$load_default_outliers, {
  tryCatch({
    # Check if default files exist (match initialization.R pattern)
    heel_file <- file.path(dataExtraFolder, "outliers_heelstrikes.csv")
    step_file <- file.path(dataExtraFolder, "outliers_steps.csv")
    
    heel_exists <- file.exists(heel_file)
    step_exists <- file.exists(step_file)
    
    if (!heel_exists && !step_exists) {
      # Detailed error message with debugging info
      debug_msg <- paste0(
        "No default outlier files found!\n\n",
        "Working directory: ", getwd(), "\n",
        "dataExtraFolder: ", dataExtraFolder, "\n\n",
        "Checked paths:\n",
        "• ", heel_file, " (exists: ", heel_exists, ")\n",
        "• ", step_file, " (exists: ", step_exists, ")\n\n",
        "Available files in data_extra folder:\n",
        paste(list.files(dataExtraFolder, full.names = FALSE), collapse = ", ")
      )
      showNotification(debug_msg, type = "error", duration = 15)
      return()
    }
    
    # Track loading results
    heel_loaded <- 0
    step_loaded <- 0
    errors <- c()
    
    # Load heel strike outliers if file exists
    if (heel_exists) {
      tryCatch({
        df_heel <- read.csv(heel_file, stringsAsFactors = FALSE)
        
        # Validate required columns
        if (all(c("participant", "trialNum", "time") %in% colnames(df_heel))) {
          # Clean and standardize data
          df_heel$trialNum <- as.numeric(df_heel$trialNum)
          df_heel$participant <- as.character(df_heel$participant)
          df_heel <- df_heel[, c("participant", "trialNum", "time"), drop = FALSE]
          
          # Remove any invalid rows
          df_heel <- df_heel[!is.na(df_heel$trialNum) & !is.na(df_heel$time), ]
          
          heelstrike_outliers(df_heel)
          heel_loaded <- nrow(df_heel)
        } else {
          errors <- c(errors, "Invalid heel strike file format")
        }
      }, error = function(e) {
        errors <- c(errors, paste("Heel strike file error:", e$message))
      })
    }
    
    # Load step outliers if file exists  
    if (step_exists) {
      tryCatch({
        df_step <- read.csv(step_file, stringsAsFactors = FALSE)
        
        # Validate required columns (support both old and new format)
        if (all(c("participant", "trialNum", "time") %in% colnames(df_step))) {
          # New simplified format
          df_step$trialNum <- as.numeric(df_step$trialNum)
          df_step$participant <- as.character(df_step$participant)
          df_step <- df_step[, c("participant", "trialNum", "time"), drop = FALSE]
        } else if (all(c("participant", "trialNum", "time_start", "time_end") %in% colnames(df_step))) {
          # Old format - convert to new format using start time
          df_step$time <- df_step$time_start
          df_step$trialNum <- as.numeric(df_step$trialNum)
          df_step$participant <- as.character(df_step$participant)
          df_step <- df_step[, c("participant", "trialNum", "time"), drop = FALSE]
        } else {
          errors <- c(errors, "Invalid step outlier file format")
          df_step <- NULL
        }
        
        if (!is.null(df_step)) {
          # Remove any invalid rows
          df_step <- df_step[!is.na(df_step$trialNum) & !is.na(df_step$time), ]
          
          step_outliers(df_step)
          step_loaded <- nrow(df_step)
        }
      }, error = function(e) {
        errors <- c(errors, paste("Step outlier file error:", e$message))
      })
    }
    
    # Show results notification
    if (length(errors) > 0) {
      showNotification(paste("Errors occurred:", paste(errors, collapse = "; ")), type = "error", duration = 10)
    } else {
      success_msg <- paste0(
        "✅ Default outlier files loaded successfully!\n",
        "❌ Heel strike outliers: ", heel_loaded, " entries\n",
        "◐ Step outliers: ", step_loaded, " entries"
      )
      showNotification(success_msg, type = "message", duration = 5)
    }
    
  }, error = function(e) {
    showNotification(paste("Error loading default outlier files:", e$message), type = "error", duration = 10)
  })
})

# Reactive values to store heel strike and step outliers separately
heelstrike_outliers <- reactiveVal(data.frame(
  participant = character(0),
  trialNum = numeric(0),
  time = numeric(0),
  stringsAsFactors = FALSE
))

step_outliers <- reactiveVal(data.frame(
  participant = character(0),
  trialNum = numeric(0),
  time = numeric(0),
  stringsAsFactors = FALSE
))

# Current outliers for the selected participant and trial
current_heelstrike_outliers <- reactive({
  heelstrike_outliers() %>%
    dplyr::filter(participant == input$participant & trialNum == as.numeric(input$trialNum))
})

current_step_outliers <- reactive({
  step_outliers() %>%
    dplyr::filter(participant == input$participant & trialNum == as.numeric(input$trialNum))
})

# Reactive values to store current axis limits (for plot memory)
axis_limits <- reactiveValues(x = NULL, y = NULL)



# Simple plot refresh observer
observeEvent(input$refresh_plot, {
  # Show warning if there are many outliers and auto-disable if too many
  total_outliers <- 0
  if (input$show_outliers_on_plot) {
    num_hs_outliers <- nrow(heelstrike_outliers())
    num_step_outliers <- nrow(step_outliers())
    total_outliers <- num_hs_outliers + num_step_outliers
    
    if (total_outliers > 500) {
      updateCheckboxInput(session, "show_outliers_on_plot", value = FALSE)
      showNotification(paste("Auto-disabled outlier display due to", total_outliers, "outliers. Enable manually if needed."), 
                       type = "warning", duration = 5)
    } else if (total_outliers > 100) {
      showNotification(paste("Refreshing plot with", total_outliers, "outliers - this may take a moment..."), 
                       type = "message", duration = 3)
    }
  }
  
  refresh_trigger(refresh_trigger() + 1)
})

# Observe changes in the plot's layout to remember axis limits
observeEvent(event_data("plotly_relayout"), {
  relayout_data <- event_data("plotly_relayout")
  
  if (!is.null(relayout_data)) {
    # Update x-axis limits
    if (!is.null(relayout_data[["xaxis.range[0]"]]) && !is.null(relayout_data[["xaxis.range[1]"]])) {
      axis_limits$x <- c(relayout_data[["xaxis.range[0]"]], relayout_data[["xaxis.range[1]"]])
    } else if (!is.null(relayout_data$xaxis.range)) {
      axis_limits$x <- relayout_data$xaxis.range
    }

    # Update y-axis limits
    if (!is.null(relayout_data[["yaxis.range[0]"]]) && !is.null(relayout_data[["yaxis.range[1]"]])) {
      axis_limits$y <- c(relayout_data[["yaxis.range[0]"]], relayout_data[["yaxis.range[1]"]])
    } else if (!is.null(relayout_data$yaxis.range)) {
      axis_limits$y <- relayout_data$yaxis.range
    }
  }
})

# Update outliers based on plot clicks (mode-dependent)
observeEvent(event_data("plotly_click"), {
  click_data <- event_data("plotly_click")
  
  if (!is.null(click_data)) {
    current_participant <- input$participant
    current_trial <- as.numeric(input$trialNum)
    clicked_time <- click_data$x
    
    params <- filteredParams()
    trial_data <- params[params$participant == current_participant &
                        params$trialNum == current_trial, ]
    
    if (nrow(trial_data) == 0) return()
    
    if (input$outlier_mode == "step") {
      # Step mode: find closest heel strike to mark as step outlier
      closest_step <- trial_data %>%
        mutate(time_diff = abs(time - clicked_time)) %>%
        arrange(time_diff) %>%
        slice(1)
      
      if (nrow(closest_step) > 0 && closest_step$time_diff <= 0.25) {
        current_step_outliers <- step_outliers()
        current_hs_outliers <- heelstrike_outliers()
        
        # Check if this heel strike is already a heel strike outlier
        existing_hs_outlier <- current_hs_outliers %>%
          dplyr::filter(participant == current_participant,
                 trialNum == current_trial,
                 abs(time - closest_step$time) <= input$outlier_threshold)
        
        if (nrow(existing_hs_outlier) > 0) {
          # Move from heel strike outlier to step outlier
          # Remove from heel strike outliers
          updated_hs_outliers <- current_hs_outliers %>%
            dplyr::filter(!(participant == current_participant &
                     trialNum == current_trial &
                     abs(time - closest_step$time) <= input$outlier_threshold))
          heelstrike_outliers(updated_hs_outliers)
          
          # Add to step outliers
          new_step_outlier <- data.frame(
            participant = current_participant,
            trialNum    = current_trial,
            time        = closest_step$time,
            stringsAsFactors = FALSE
          )
          updated_step_outliers <- rbind(current_step_outliers, new_step_outlier)
          step_outliers(updated_step_outliers)
          showNotification("Moved from heel strike removal to step outlier", type = "message")
        } else {
          # Check if this step already exists as step outlier
          existing_step <- current_step_outliers %>%
            dplyr::filter(participant == current_participant,
                   trialNum == current_trial,
                   abs(time - closest_step$time) <= input$outlier_threshold)
          
          if (nrow(existing_step) > 0) {
            # Remove the step outlier
            updated_step_outliers <- current_step_outliers %>%
              dplyr::filter(!(participant == current_participant &
                       trialNum == current_trial &
                       abs(time - closest_step$time) <= input$outlier_threshold))
            step_outliers(updated_step_outliers)
            showNotification("Step outlier removed", type = "message")
          } else {
            # Add the step outlier
            new_step_outlier <- data.frame(
              participant = current_participant,
              trialNum    = current_trial,
              time        = closest_step$time,
              stringsAsFactors = FALSE
            )
            updated_step_outliers <- rbind(current_step_outliers, new_step_outlier)
            step_outliers(updated_step_outliers)
            showNotification("Step outlier added", type = "message")
          }
        }
      }
    } else {
      # Heel strike mode: find closest heel strike
      closest_step <- trial_data %>%
        mutate(time_diff = abs(time - clicked_time)) %>%
        arrange(time_diff) %>%
        slice(1)
      
      if (nrow(closest_step) > 0 && closest_step$time_diff <= 0.25) {
        current_hs_outliers <- heelstrike_outliers()
        current_step_outliers <- step_outliers()
        
        # Check if this heel strike is already a step outlier
        existing_step_outlier <- current_step_outliers %>%
          dplyr::filter(participant == closest_step$participant,
                 trialNum == closest_step$trialNum,
                 abs(time - closest_step$time) <= input$outlier_threshold)
        
        if (nrow(existing_step_outlier) > 0) {
          # Move from step outlier to heel strike outlier
          # Remove from step outliers
          updated_step_outliers <- current_step_outliers %>%
            dplyr::filter(!(participant == closest_step$participant &
                     trialNum == closest_step$trialNum &
                     abs(time - closest_step$time) <= input$outlier_threshold))
          step_outliers(updated_step_outliers)
          
          # Add to heel strike outliers
          new_outlier <- data.frame(
            participant = closest_step$participant,
            trialNum    = closest_step$trialNum,
            time        = closest_step$time,
            stringsAsFactors = FALSE
          )
          updated_outliers <- rbind(current_hs_outliers, new_outlier)
          heelstrike_outliers(updated_outliers)
          showNotification("Moved from step outlier to heel strike removal", type = "message")
        } else {
          # Check if this heel strike already exists as heel strike outlier
          existing_outlier <- current_hs_outliers %>%
            dplyr::filter(participant == closest_step$participant,
                   trialNum == closest_step$trialNum,
                   abs(time - closest_step$time) <= input$outlier_threshold)
          
          if (nrow(existing_outlier) > 0) {
            # Remove the heel strike outlier
            updated_outliers <- current_hs_outliers %>%
              dplyr::filter(!(participant == closest_step$participant &
                       trialNum == closest_step$trialNum &
                       abs(time - closest_step$time) <= input$outlier_threshold))
            heelstrike_outliers(updated_outliers)
            showNotification("Heel strike outlier removed", type = "message")
          } else {
            # Add the heel strike outlier
            new_outlier <- data.frame(
              participant = closest_step$participant,
              trialNum    = closest_step$trialNum,
              time        = closest_step$time,
              stringsAsFactors = FALSE
            )
            updated_outliers <- rbind(current_hs_outliers, new_outlier)
            heelstrike_outliers(updated_outliers)
            showNotification("Heel strike outlier added", type = "message")
          }
        }
      }
    }
  }
})

# Navigation functionality for special heel strikes
special_heel_strikes <- reactive({
  req(filteredParams(), input$participant, input$trialNum)
  
  current_participant <- input$participant
  current_trial <- as.numeric(input$trialNum)
  
  # Get current trial data
  trial_data <- filteredParams() %>%
    dplyr::filter(participant == current_participant & trialNum == current_trial) %>%
    dplyr::arrange(time)
  
  if (nrow(trial_data) == 0) {
    return(data.frame(time = numeric(0), type = character(0), stringsAsFactors = FALSE))
  }
  
  special_times <- c()
  special_types <- c()
  
  # Add suspect steps
  if ("suspect" %in% colnames(trial_data)) {
    suspect_steps <- trial_data %>% dplyr::filter(suspect == TRUE)
    if (nrow(suspect_steps) > 0) {
      special_times <- c(special_times, suspect_steps$time)
      special_types <- c(special_types, rep("suspect", nrow(suspect_steps)))
    }
  }
  
  # Add heel strike outliers
  hs_outliers <- current_heelstrike_outliers()
  if (nrow(hs_outliers) > 0) {
    special_times <- c(special_times, hs_outliers$time)
    special_types <- c(special_types, rep("heel_strike_outlier", nrow(hs_outliers)))
  }
  
  # Add step outliers
  step_outliers_data <- current_step_outliers()
  if (nrow(step_outliers_data) > 0) {
    special_times <- c(special_times, step_outliers_data$time)
    special_types <- c(special_types, rep("step_outlier", nrow(step_outliers_data)))
  }
  
  # Create combined data frame and sort by time
  if (length(special_times) > 0) {
    special_df <- data.frame(
      time = special_times,
      type = special_types,
      stringsAsFactors = FALSE
    ) %>%
      dplyr::arrange(time) %>%
      dplyr::distinct(time, .keep_all = TRUE)  # Remove duplicates, keeping first type
  } else {
    special_df <- data.frame(time = numeric(0), type = character(0), stringsAsFactors = FALSE)
  }
  
  return(special_df)
})

# Function to navigate to a specific time using Plotly proxy (fast viewport update)
navigate_to_time <- function(target_time, current_x_axis, current_y_axis) {
  # Get current axis ranges or use data ranges if not set
  current_data <- filteredParams() %>%
    dplyr::filter(participant == input$participant & trialNum == as.numeric(input$trialNum))
  
  if (nrow(current_data) == 0) return()
  
  # Calculate current window sizes
  if (!is.null(axis_limits$x) && length(axis_limits$x) == 2) {
    x_window_size <- axis_limits$x[2] - axis_limits$x[1]
  } else {
    # Default window size (use 10% of data range)
    x_range <- range(current_data[[current_x_axis]], na.rm = TRUE)
    x_window_size <- (x_range[2] - x_range[1]) * 0.1
  }
  
  if (!is.null(axis_limits$y) && length(axis_limits$y) == 2) {
    y_window_size <- axis_limits$y[2] - axis_limits$y[1]
  } else {
    # Default window size (use 10% of data range)
    y_range <- range(current_data[[current_y_axis]], na.rm = TRUE)
    y_window_size <- (y_range[2] - y_range[1]) * 0.1
  }
  
  # Center the windows on the target time
  x_center <- target_time
  y_center <- current_data %>%
    dplyr::filter(abs(time - target_time) == min(abs(time - target_time))) %>%
    dplyr::slice(1) %>%
    dplyr::pull(!!sym(current_y_axis))
  
  if (length(y_center) == 0) {
    y_center <- mean(current_data[[current_y_axis]], na.rm = TRUE)
  }
  
  # Calculate new axis limits
  new_x_range <- c(x_center - x_window_size/2, x_center + x_window_size/2)
  new_y_range <- c(y_center - y_window_size/2, y_center + y_window_size/2)
  
  # Update axis limits in memory for future reference
  axis_limits$x <- new_x_range
  axis_limits$y <- new_y_range
  
  # Use plotlyProxy to update just the viewport (much faster than full refresh)
  plotlyProxy("manualOutlierPlot", session) %>%
    plotlyProxyInvoke("relayout", list(
      "xaxis.range" = new_x_range,
      "yaxis.range" = new_y_range
    ))
}

# Previous special heel strike navigation
observeEvent(input$nav_prev_special, {
  req(input$xplot_16, input$yplot_16)
  
  special_data <- special_heel_strikes()
  if (nrow(special_data) == 0) {
    showNotification("No special heel strikes found for this participant/trial", type = "message")
    return()
  }
  
  # Get current center time
  current_center_time <- if (!is.null(axis_limits$x) && length(axis_limits$x) == 2) {
    mean(axis_limits$x)
  } else {
    # If no zoom set, use middle of current trial data
    trial_data <- filteredParams() %>%
      dplyr::filter(participant == input$participant & trialNum == as.numeric(input$trialNum))
    if (nrow(trial_data) > 0) {
      mean(range(trial_data$time, na.rm = TRUE))
    } else {
      return()
    }
  }
  
  # Find previous special heel strike (at least threshold distance away)
  min_threshold <- input$outlier_threshold
  prev_specials <- special_data %>% 
    dplyr::filter(time < (current_center_time - min_threshold)) %>% 
    dplyr::arrange(desc(time))
  
  if (nrow(prev_specials) > 0) {
    target_time <- prev_specials$time[1]
    target_type <- prev_specials$type[1]
    navigate_to_time(target_time, input$xplot_16, input$yplot_16)
    showNotification(paste("Navigated to previous", gsub("_", " ", target_type), "at time", round(target_time, 3), 
                          "(≥", min_threshold, "s away)"), type = "message")
  } else {
    # Wrap around to last special heel strike
    target_time <- special_data$time[nrow(special_data)]
    target_type <- special_data$type[nrow(special_data)]
    navigate_to_time(target_time, input$xplot_16, input$yplot_16)
    showNotification(paste("Wrapped to last", gsub("_", " ", target_type), "at time", round(target_time, 3)), type = "message")
  }
})

# Next special heel strike navigation
observeEvent(input$nav_next_special, {
  req(input$xplot_16, input$yplot_16)
  
  special_data <- special_heel_strikes()
  if (nrow(special_data) == 0) {
    showNotification("No special heel strikes found for this participant/trial", type = "message")
    return()
  }
  
  # Get current center time
  current_center_time <- if (!is.null(axis_limits$x) && length(axis_limits$x) == 2) {
    mean(axis_limits$x)
  } else {
    # If no zoom set, use middle of current trial data
    trial_data <- filteredParams() %>%
      dplyr::filter(participant == input$participant & trialNum == as.numeric(input$trialNum))
    if (nrow(trial_data) > 0) {
      mean(range(trial_data$time, na.rm = TRUE))
    } else {
      return()
    }
  }
  
  # Find next special heel strike (at least threshold distance away)
  min_threshold <- input$outlier_threshold
  next_specials <- special_data %>% 
    dplyr::filter(time > (current_center_time + min_threshold)) %>% 
    dplyr::arrange(time)
  
  if (nrow(next_specials) > 0) {
    target_time <- next_specials$time[1]
    target_type <- next_specials$type[1]
    navigate_to_time(target_time, input$xplot_16, input$yplot_16)
    showNotification(paste("Navigated to next", gsub("_", " ", target_type), "at time", round(target_time, 3), 
                          "(≥", min_threshold, "s away)"), type = "message")
  } else {
    # Wrap around to first special heel strike
    target_time <- special_data$time[1]
    target_type <- special_data$type[1]
    navigate_to_time(target_time, input$xplot_16, input$yplot_16)
    showNotification(paste("Wrapped to first", gsub("_", " ", target_type), "at time", round(target_time, 3)), type = "message")
  }
})

# Status text showing current special heel strikes count
output$nav_status <- renderText({
  special_data <- special_heel_strikes()
  if (nrow(special_data) == 0) {
    "No special heel strikes found"
  } else {
    paste("Found", nrow(special_data), "special heel strikes:",
          sum(special_data$type == "suspect"), "suspect,",
          sum(special_data$type == "heel_strike_outlier"), "heel strike outliers,", 
          sum(special_data$type == "step_outlier"), "step outliers")
  }
})

# Optimize the plot rendering with caching
plot_cache <- reactiveVal(NULL)
plot_participant <- reactiveVal(NULL)
plot_trial <- reactiveVal(NULL)
plot_xaxis <- reactiveVal(NULL)
plot_yaxis <- reactiveVal(NULL)

# Render the interactive plot
output$manualOutlierPlot <- renderPlotly({
  
  # Check if we need to regenerate the base plot
  current_participant <- input$participant
  current_trial <- as.numeric(input$trialNum)
  current_xaxis <- input$xplot_16
  current_yaxis <- input$yplot_16
  
  # Only regenerate base plot if parameters changed or forced refresh
  current_refresh_trigger <- refresh_trigger()
  need_new_plot <- is.null(plot_cache()) || 
                   plot_participant() != current_participant || 
                   plot_trial() != current_trial ||
                   plot_xaxis() != current_xaxis ||
                   plot_yaxis() != current_yaxis ||
                   current_refresh_trigger != last_refresh_trigger()  # Only if refresh was triggered
  
  if (need_new_plot) {
    # Check if we have data to plot
    plot_data <- filteredParams()
    if (is.null(plot_data) || nrow(plot_data) == 0) {
      return(plot_ly() %>% 
        layout(title = "No data available for the selected participant/trial"))
    }
    
    # Create the base ggplot
    base_plot <- tryCatch({
      plot_steps(
        plot_data, current_participant, current_trial,
        current_xaxis, current_yaxis, input$filter, 
        baseSize = input$baseSize
      )
    }, error = function(e) {
      warning("Error creating base plot: ", e$message)
      return(plot_ly() %>% layout(title = paste("Error creating plot:", e$message)))
    })
    
    # Cache the base plot and parameters
    plot_cache(base_plot)
    plot_participant(current_participant)
    plot_trial(current_trial)  
    plot_xaxis(current_xaxis)
    plot_yaxis(current_yaxis)
    last_refresh_trigger(current_refresh_trigger)  # Update the last trigger value
  } else {
    # Use cached base plot
    base_plot <- plot_cache()
  }
  
  # Ensure base plot exists
  if (is.null(base_plot)) {
    return(plot_ly() %>% 
      layout(title = "Unable to generate plot"))
  }

  # Always update outlier highlighting (this is fast)
  req(filteredParams())
  
  # Get the current data for this participant and trial
  current_data <- filteredParams() %>%
    dplyr::filter(participant == current_participant & trialNum == current_trial)
  
  # Add comprehensive hover text for all heel strikes
  if (nrow(current_data) > 0) {
    current_data$plot_text <- create_heel_strike_hover_text(current_data, current_xaxis, current_yaxis, "normal")
  }
  
  # Only process outliers if they need to be shown on plot (performance optimization)
  current_data$is_hs_outlier <- FALSE
  current_hs_outliers <- NULL
  current_step_outliers_data <- NULL
  
  if (input$show_outliers_on_plot) {
    # Get current outliers for both types
    current_hs_outliers <- current_heelstrike_outliers()
    current_step_outliers_data <- current_step_outliers()
    
    # Mark heel strike outliers efficiently
    if (nrow(current_hs_outliers) > 0 && nrow(current_data) > 0) {
      threshold <- input$outlier_threshold
      
      for (i in seq_len(nrow(current_hs_outliers))) {
        within_range <- abs(current_data$time - current_hs_outliers$time[i]) <= threshold
        current_data$is_hs_outlier <- current_data$is_hs_outlier | within_range
      }
    }
  }
  
  # suspect column present
  if (!"suspect" %in% colnames(current_data)) current_data$suspect <- FALSE
  
  # Add outlier highlighting to the plot
  p <- base_plot
  
  # Add invisible overlay for comprehensive hover text on all heel strikes
  if (nrow(current_data) > 0) {
    p <- p + geom_point(data = current_data,
                       aes(x = .data[[current_xaxis]],
                           y = .data[[current_yaxis]],
                           text = plot_text),
                       alpha = 0, size = 0.1)  # Invisible but with hover text
  }
  
    # First add suspect points (orange)
  suspect_data <- current_data %>% dplyr::filter(suspect)
  if (nrow(suspect_data) > 0) {
    # Enhance hover text for suspect points
    suspect_data$plot_text <- create_heel_strike_hover_text(suspect_data, current_xaxis, current_yaxis, "suspect", "", "Flagged as suspect")
    
    p <- p + geom_point(data = suspect_data,
                       aes(x = .data[[current_xaxis]],
                           y = .data[[current_yaxis]],
                           text = plot_text),
                       color = "orange", size = 4, shape = 21, fill = "orange", alpha = 0.6)
  }
  

  
    # Add heel strike outliers (X symbol, yellow)
  hs_outlier_data <- current_data %>% dplyr::filter(is_hs_outlier)
  if (nrow(hs_outlier_data) > 0) {
    # Enhanced hover text for heel strike outliers
    hs_outlier_data$plot_text <- create_heel_strike_hover_text(hs_outlier_data, current_xaxis, current_yaxis, "heel_outlier", "", "Marked for heel strike removal")
    
    p <- p + geom_point(data = hs_outlier_data,
                         aes(x = .data[[current_xaxis]],
                             y = .data[[current_yaxis]],
                             text = plot_text),
                       color = "gold", size = 4, shape = 4, alpha = 0.8)  # Shape 4 = X, yellow color
  }
  
  # Add step outlier markers (left half circle, grey) - mark heel strikes for step outliers
  if (input$show_outliers_on_plot && !is.null(current_step_outliers_data) && nrow(current_step_outliers_data) > 0) {
    # Create markers at the heel strike times for step outliers
    step_marker_data <- current_data %>%
      dplyr::filter(
        sapply(time, function(t) {
          any(abs(current_step_outliers_data$time - t) <= input$outlier_threshold)
        })
      )
    
    if (nrow(step_marker_data) > 0) {
      # Enhanced hover text for step outliers
      step_marker_data$plot_text <- create_heel_strike_hover_text(step_marker_data, current_xaxis, current_yaxis, "step_outlier", "", "Marked as step outlier for analysis")
      
      p <- p + geom_text(data = step_marker_data,
                         aes(x = .data[[current_xaxis]],
                             y = .data[[current_yaxis]],
                             text = plot_text),
                         label = "◐", color = "grey60", size = 12, alpha = 0.6)  # Unicode left half circle, 2x larger
    }
  }
  
  # Get current axis limits (if they exist from previous zoom/pan)
  x_range_current <- isolate(axis_limits$x)
  y_range_current <- isolate(axis_limits$y)
  
  # Convert to plotly with interactive zoom and pan
  p <- ggplotly(p, width = input$plotwidth, height = input$plotheight, tooltip = "text") %>%
    style(
      hoverinfo = "none",
      traces = which(sapply(.$x$data, function(d) d$type) == "scatter" & sapply(.$x$data, function(d) d$mode) == "lines")
    ) %>%
    layout(
      dragmode = "pan"
    ) %>%
    event_register("plotly_click") %>%
    event_register("plotly_relayout")
  
  # Apply remembered axis limits if they exist
  if (!is.null(x_range_current) && !is.null(y_range_current)) {
    p <- p %>%
      layout(
        xaxis = list(range = x_range_current),
        yaxis = list(range = y_range_current)
      )
  }
  
  return(p)
})



# Render the heel strike outliers table
output$heelStrikeOutliersTable <- DT::renderDataTable({
  data <- if (input$filter_outliers_table) {
    current_heelstrike_outliers()
  } else {
    heelstrike_outliers()
  }
  
  result <- create_outlier_table(data, "heelStrikeOutliersTable", "heelstrike", heelstrike_table_initialized)
  
  # Ensure we always return a table
  if (is.null(result)) {
    empty_data <- data.frame(
      Message = "No heel strike outliers found. Click on plot to add outliers.",
      stringsAsFactors = FALSE
    )
    return(DT::datatable(
      empty_data,
      options = list(dom = 't', ordering = FALSE, searching = FALSE, 
                     paging = FALSE, info = FALSE),
      selection = 'none', escape = FALSE
    ))
  }
  
  return(result)
})

# Render the step outliers table  
output$stepOutliersTable <- DT::renderDataTable({
  data <- if (input$filter_outliers_table) {
    current_step_outliers()
  } else {
    step_outliers()
  }
  
  result <- create_outlier_table(data, "stepOutliersTable", "step", step_table_initialized)
  
  # Ensure we always return a table
  if (is.null(result)) {
    empty_data <- data.frame(
      Message = "No step outliers found. Click on plot to add outliers.",
      stringsAsFactors = FALSE
    )
    return(DT::datatable(
      empty_data,
      options = list(dom = 't', ordering = FALSE, searching = FALSE, 
                     paging = FALSE, info = FALSE),
      selection = 'none', escape = FALSE
    ))
  }
  
  return(result)
})

# Track if tables have been initialized to decide between full render vs proxy update
heelstrike_table_initialized <- reactiveVal(FALSE)
step_table_initialized <- reactiveVal(FALSE)

# Reusable function to create outlier tables (optimized for performance)
create_outlier_table <- function(data, table_id, table_type, initialized_flag) {
  current_participant <- input$participant
  current_trial <- as.numeric(input$trialNum)
  
  # Track filter mode changes (only if needed)
  current_filter_mode <- input$filter_outliers_table
  filter_var_name <- paste0("last_filter_mode_", table_type)
  
  mode_changed <- FALSE
  if (!exists(filter_var_name, envir = .GlobalEnv) || 
      get(filter_var_name, envir = .GlobalEnv) != current_filter_mode) {
    assign(filter_var_name, current_filter_mode, envir = .GlobalEnv)
    initialized_flag(FALSE)
    mode_changed <- TRUE
  }
  
  # Handle empty data case efficiently
  if (is.null(data) || nrow(data) == 0) {
    empty_data <- data.frame(
      Message = paste("No", table_type, "outliers found. Click on plot to add outliers."),
      stringsAsFactors = FALSE
    )
    
    initialized_flag(TRUE)  # Mark as initialized
    return(DT::datatable(
      empty_data,
      options = list(dom = 't', ordering = FALSE, searching = FALSE, 
                     paging = FALSE, info = FALSE),
      selection = 'none', escape = FALSE
    ))
  }
  
  # Lightweight change detection - only check row count and mode changes
  simple_hash <- paste(nrow(data), current_filter_mode, current_participant, current_trial, sep = "_")
  
  hash_var_name <- paste0("table_simple_hash_", table_type)
  previous_hash <- if(exists(hash_var_name, envir = .GlobalEnv)) {
    get(hash_var_name, envir = .GlobalEnv)
  } else NULL
  
  assign(hash_var_name, simple_hash, envir = .GlobalEnv)
  needs_full_rerender <- is.null(previous_hash) || previous_hash != simple_hash
  
  # Skip re-render if no changes - but still return the existing table
  # For now, let's always render to ensure tables show up properly
  # if (initialized_flag() && !needs_full_rerender && !mode_changed) {
  #   return(NULL)
  # }
  
  # Prepare data for display
  if (table_type == "heelstrike") {
    display_data <- data %>%
      dplyr::mutate(
        Time = round(time, 4),
        unique_id = paste0("del_hs_", gsub("[^A-Za-z0-9]", "_", participant), "_", trialNum, "_", round(time * 1000)),
        Delete = paste0('<button class="btn btn-xs btn-danger" onclick="Shiny.setInputValue(\'', 
                       unique_id, '\', true, {priority: \'event\'})">Delete</button>')
      ) %>%
      dplyr::select(Participant = participant, Trial = trialNum, Time, Delete)
  } else {
    display_data <- data %>%
      dplyr::mutate(
        Time = round(time, 4),
        unique_id = paste0("del_step_", gsub("[^A-Za-z0-9]", "_", participant), "_", trialNum, "_", 
                          round(time * 1000)),
        Delete = paste0('<button class="btn btn-xs btn-danger" onclick="Shiny.setInputValue(\'', 
                       unique_id, '\', true, {priority: \'event\'})">Delete</button>')
      ) %>%
      dplyr::select(Participant = participant, Trial = trialNum, Time, Delete)
  }
  
  # Standard table background
  bg_color <- "rgba(245, 245, 245, 0.3)"
  
  # Create DataTable
  dt_table <- DT::datatable(
    display_data,
    options = list(
      pageLength = 50,
      stateSave = TRUE,
      stateDuration = 60 * 60 * 24,
      order = list(list(2, 'desc')), # Sort by time
      dom = 'frtip',
      columnDefs = list(list(targets = ncol(display_data) - 1, orderable = FALSE))
    ),
    selection = 'none',
    escape = FALSE,
    elementId = table_id
  ) %>%
    DT::formatStyle(columns = colnames(display_data), backgroundColor = bg_color)
  
  initialized_flag(TRUE)
  return(dt_table)
}



# Manual backup functionality
observeEvent(input$manual_backup, {
  # Check if there are any outliers to save
  num_heelstrike <- nrow(heelstrike_outliers())
  num_step <- nrow(step_outliers())
  
  if (num_heelstrike == 0 && num_step == 0) {
    showNotification("No outliers to backup - both tables are empty", type = "message")
    return()
  }
  
  # Create confirmation message with details
  backup_message <- paste0(
    "Manual Backup Confirmation\n\n",
    "This will overwrite the existing CSV files with current table data:\n\n",
    "• Heel Strike Outliers: ", num_heelstrike, " entries\n",
    "  → Will save to: ", file.path(dataExtraFolder, "outliers_heelstrikes.csv"), "\n\n",
    "• Step Outliers: ", num_step, " entries\n",
    "  → Will save to: ", file.path(dataExtraFolder, "outliers_steps.csv"), "\n\n",
    "⚠️ This will permanently overwrite existing files!\n\n",
    "Continue with backup?"
  )
  
  # Show confirmation dialog
  showModal(modalDialog(
    title = "💾 Manual Backup Confirmation",
    HTML(paste0(
      "<div style='font-family: monospace; white-space: pre-line; background-color: #f8f9fa; padding: 15px; border-radius: 5px; margin: 10px 0;'>",
      "<strong>Current Table Contents:</strong><br><br>",
      "❌ <strong>Heel Strike Outliers:</strong> ", num_heelstrike, " entries<br>",
      "   → File: <code>", file.path(dataExtraFolder, "outliers_heelstrikes.csv"), "</code><br><br>",
      "◐ <strong>Step Outliers:</strong> ", num_step, " entries<br>", 
      "   → File: <code>", file.path(dataExtraFolder, "outliers_steps.csv"), "</code><br><br>",
      "<span style='color: #d9534f;'><strong>⚠️ Warning:</strong> This will permanently overwrite existing files!</span>",
      "</div>"
    )),
    footer = tagList(
      modalButton("❌ Cancel"),
      actionButton("confirm_manual_backup", "✅ Confirm Backup", 
                   class = "btn-warning", style = "font-weight: bold;")
    ),
    easyClose = TRUE,
    size = "m"
  ))
})

# Handle confirmed manual backup
observeEvent(input$confirm_manual_backup, {
  removeModal()
  
  tryCatch({
    backup_success <- TRUE
    
    # Save heel strike outliers (match initialization.R pattern)
    heel_save_path <- file.path(dataExtraFolder, "outliers_heelstrikes.csv")
    if (nrow(heelstrike_outliers()) > 0) {
      write.csv(heelstrike_outliers(), heel_save_path, row.names = FALSE)
    } else {
      # Create empty file if no data
      write.csv(data.frame(participant = character(0), trialNum = numeric(0), time = numeric(0)), 
                heel_save_path, row.names = FALSE)
    }
    
    # Save step outliers (match initialization.R pattern)
    step_save_path <- file.path(dataExtraFolder, "outliers_steps.csv")
    if (nrow(step_outliers()) > 0) {
      write.csv(step_outliers(), step_save_path, row.names = FALSE)
    } else {
      # Create empty file if no data
      write.csv(data.frame(participant = character(0), trialNum = numeric(0), time = numeric(0)), 
                step_save_path, row.names = FALSE)
    }
    
    showNotification(
      paste("✅ Manual backup successful!", 
            nrow(heelstrike_outliers()), "heel strike outliers and", 
            nrow(step_outliers()), "step outliers saved to CSV files"),
      type = "message", duration = 5
    )
    
  }, error = function(e) {
    showNotification(paste("❌ Backup failed:", e$message), type = "error", duration = 10)
  })
})

# Export rotations to CSV
# Export handlers 
output$export_heelstrike_outliers <- downloadHandler(
  filename = function() {
    paste0("heelstrike_outliers_", format(Sys.time(), "%Y%m%d_%H%M%S"), ".csv")
  },
  content = function(file) {
    write.csv(heelstrike_outliers(), file, row.names = FALSE)
  }
)

output$export_step_outliers <- downloadHandler(
  filename = function() {
    paste0("step_outliers_", format(Sys.time(), "%Y%m%d_%H%M%S"), ".csv")
  },
  content = function(file) {
    write.csv(step_outliers(), file, row.names = FALSE)
  }
)

# Handle delete button events for heel strike outliers
observe({
  all_inputs <- names(input)
  del_hs_inputs <- all_inputs[grepl("^del_hs_", all_inputs)]
  
  lapply(del_hs_inputs, function(input_id) {
    observeEvent(input[[input_id]], {
      if (!is.null(input[[input_id]]) && input[[input_id]]) {
        tryCatch({
          # Extract participant, trial, and time from the input_id
          # Format: del_hs_PARTICIPANT_TRIALNUMBER_TIMEMILLISECONDS
          parts <- strsplit(input_id, "_")[[1]]
          if (length(parts) >= 4) {
            participant <- gsub("_", " ", parts[3])
            trial_num <- as.numeric(parts[4])
            time_ms <- as.numeric(parts[5])
            time_sec <- time_ms / 1000
            
            # Remove from heel strike outliers
            current_hs_outliers <- heelstrike_outliers()
            updated_outliers <- current_hs_outliers %>%
              dplyr::filter(!(participant == !!participant &
                       trialNum == !!trial_num &
                       abs(time - !!time_sec) <= !!input$outlier_threshold))
            
            heelstrike_outliers(updated_outliers)
            showNotification("Heel strike outlier deleted", type = "message", duration = 1)
          }
        }, error = function(e) {
          showNotification(paste("Error deleting heel strike outlier:", e$message), type = "error")
        })
      }
    }, ignoreInit = TRUE)
  })
})

# Handle delete button events for step outliers
observe({
  all_inputs <- names(input)
  del_step_inputs <- all_inputs[grepl("^del_step_", all_inputs)]
  
  lapply(del_step_inputs, function(input_id) {
    observeEvent(input[[input_id]], {
      if (!is.null(input[[input_id]]) && input[[input_id]]) {
        tryCatch({
          # Extract participant, trial, and time from the input_id
          # Format: del_step_PARTICIPANT_TRIALNUMBER_TIMEMILLISECONDS  
          parts <- strsplit(input_id, "_")[[1]]
          if (length(parts) >= 4) {
            participant <- gsub("_", " ", parts[3])
            trial_num <- as.numeric(parts[4])
            time_ms <- as.numeric(parts[5])
            time_sec <- time_ms / 1000
            
            # Remove from step outliers
            current_step_outliers <- step_outliers()
            updated_outliers <- current_step_outliers %>%
              dplyr::filter(!(participant == !!participant &
                       trialNum == !!trial_num &
                       abs(time - !!time_sec) <= !!input$outlier_threshold))
            
            step_outliers(updated_outliers)
            showNotification("Step outlier deleted", type = "message", duration = 1)
          }
        }, error = function(e) {
          showNotification(paste("Error deleting step outlier:", e$message), type = "error")
        })
      }
    }, ignoreInit = TRUE)
  })
})

# Observer to overwrite outliers in allGaitParams
observeEvent(input$overwrite_outliers, {
  tryCatch({
    # Get current outliers for both types
    current_hs_outliers <- heelstrike_outliers()
    current_step_outliers_data <- step_outliers()
    
    # Calculate detailed impact statistics before applying
    total_outliers <- nrow(current_hs_outliers) + nrow(current_step_outliers_data)
    
    if (total_outliers == 0) {
      showNotification("No outliers to apply", type = "message", duration = 3)
      return()
    }
    
    # Calculate heel strike impact
    hs_impact_details <- ""
    if (nrow(current_hs_outliers) > 0) {
      # Filter out invalid participants
      valid_participants <- unique(allGaitParams$participant)
      valid_hs_outliers <- current_hs_outliers[
        as.character(current_hs_outliers$participant) %in% as.character(valid_participants), 
      ]
      
      if (nrow(valid_hs_outliers) > 0) {
        # Calculate how many heel strikes will be affected per participant/trial
        hs_summary <- data.frame()
        total_hs_affected <- 0
        
        for (i in seq_len(nrow(valid_hs_outliers))) {
          hs_outlier <- valid_hs_outliers[i, ]
          
          # Find matching heel strikes
          matches <- which(
            abs(allGaitParams$time - hs_outlier$time) <= input$outlier_threshold &
            as.character(allGaitParams$participant) == as.character(hs_outlier$participant) &
            as.numeric(allGaitParams$trialNum) == as.numeric(hs_outlier$trialNum)
          )
          
          if (length(matches) > 0) {
            total_hs_affected <- total_hs_affected + length(matches)
            participant_trial <- paste0("P", hs_outlier$participant, "_T", hs_outlier$trialNum)
            existing_row <- which(hs_summary$participant_trial == participant_trial)
            
            if (length(existing_row) > 0) {
              hs_summary$count[existing_row] <- hs_summary$count[existing_row] + length(matches)
            } else {
              hs_summary <- rbind(hs_summary, data.frame(
                participant_trial = participant_trial,
                count = length(matches),
                stringsAsFactors = FALSE
              ))
            }
          }
        }
        
        if (nrow(hs_summary) > 0) {
          hs_details <- paste(hs_summary$participant_trial, "(", hs_summary$count, ")", collapse = ", ")
          hs_impact_details <- paste0("❌ HEEL STRIKES TO REMOVE: ", total_hs_affected, " total\n   Per trial: ", hs_details, "\n\n")
        }
      }
    }
    
    # Calculate step outlier impact
    step_impact_details <- ""
    if (nrow(current_step_outliers_data) > 0) {
      # Filter out invalid participants
      valid_participants <- unique(allGaitParams$participant)
      valid_step_outliers <- current_step_outliers_data[
        as.character(current_step_outliers_data$participant) %in% as.character(valid_participants), 
      ]
      
      if (nrow(valid_step_outliers) > 0) {
        # Calculate how many data points will be affected per participant/trial
        step_summary <- data.frame()
        total_steps_affected <- 0
        
        for (i in seq_len(nrow(valid_step_outliers))) {
          step_outlier <- valid_step_outliers[i, ]
          
          # Find matching heel strikes for step outliers
          step_matches <- which(
            abs(allGaitParams$time - step_outlier$time) <= 0.01 &
            as.character(allGaitParams$participant) == as.character(step_outlier$participant) &
            as.numeric(allGaitParams$trialNum) == as.numeric(step_outlier$trialNum)
          )
          
          if (length(step_matches) > 0) {
            total_steps_affected <- total_steps_affected + length(step_matches)
            participant_trial <- paste0("P", step_outlier$participant, "_T", step_outlier$trialNum)
            existing_row <- which(step_summary$participant_trial == participant_trial)
            
            if (length(existing_row) > 0) {
              step_summary$count[existing_row] <- step_summary$count[existing_row] + length(step_matches)
            } else {
              step_summary <- rbind(step_summary, data.frame(
                participant_trial = participant_trial,
                count = length(step_matches),
                stringsAsFactors = FALSE
              ))
            }
          }
        }
        
        if (nrow(step_summary) > 0) {
          step_details <- paste(step_summary$participant_trial, "(", step_summary$count, ")", collapse = ", ")
          step_impact_details <- paste0("◐ STEP OUTLIERS TO MARK: ", total_steps_affected, " total\n   Per trial: ", step_details, "\n\n")
        }
      }
    }
    
    # Create confirmation message
    confirmation_msg <- paste0(
      "OUTLIER APPLICATION CONFIRMATION\n\n",
      hs_impact_details,
      step_impact_details,
      "⚠️  This will modify your dataset. Are you sure you want to proceed?\n\n",
      "Click OK to apply these changes or Cancel to abort."
    )
    
    # Show confirmation dialog
    showModal(modalDialog(
      title = "⚠️ Confirm Outlier Application",
      HTML(paste0("<pre style='font-family: monospace; white-space: pre-wrap; font-size: 12px;'>", 
                  confirmation_msg, "</pre>")),
      footer = tagList(
        actionButton("confirm_apply_outliers", "✅ Apply Changes", class = "btn-success"),
        modalButton("❌ Cancel")
      ),
      easyClose = FALSE,
      size = "l"
    ))
    
  }, error = function(e) {
    showNotification(paste("Error preparing outlier application:", e$message), type = "error", duration = 5)
    warning("Error in overwrite_outliers preparation: ", e$message)
  })
})

# Handle confirmed outlier application
observeEvent(input$confirm_apply_outliers, {
  tryCatch({
    # Close the modal first
    removeModal()
    
    # Get current outliers for both types
    current_hs_outliers <- heelstrike_outliers()
    current_step_outliers_data <- step_outliers()
    
    # Show progress
    total_outliers <- nrow(current_hs_outliers) + nrow(current_step_outliers_data)
    showNotification(paste("Applying", total_outliers, "outliers..."), 
                     type = "message", duration = 2)
    
    # Step 1: Remove heel strikes marked for removal
    # Initialize heelStrikeBad column if it doesn't exist
    if (!"heelStrikeBad" %in% colnames(allGaitParams)) {
      allGaitParams$heelStrikeBad <<- FALSE
    } else {
      allGaitParams$heelStrikeBad <<- FALSE
    }
    
    # Mark heel strikes for removal
    if (nrow(current_hs_outliers) > 0) {
      # Filter out invalid participants
      valid_participants <- unique(allGaitParams$participant)
      current_hs_outliers <- current_hs_outliers[
        as.character(current_hs_outliers$participant) %in% as.character(valid_participants), 
      ]
      
      for (i in seq_len(nrow(current_hs_outliers))) {
        hs_outlier <- current_hs_outliers[i, ]
        
        # Find matching heel strikes
        matches <- which(
          abs(allGaitParams$time - hs_outlier$time) <= input$outlier_threshold &
          as.character(allGaitParams$participant) == as.character(hs_outlier$participant) &
          as.numeric(allGaitParams$trialNum) == as.numeric(hs_outlier$trialNum)
        )
        
        if (length(matches) > 0) {
          allGaitParams$heelStrikeBad[matches] <<- TRUE
        }
      }
      
      n_heel_bad <- sum(allGaitParams$heelStrikeBad, na.rm = TRUE)
      showNotification(paste("Marked", n_heel_bad, "heel strikes for removal"), 
                       type = "message", duration = 1)
    }
    
    # Step 2: Reset outlier steps and then mark step outliers
    allGaitParams$outlierSteps <<- FALSE
    
    if (nrow(current_step_outliers_data) > 0) {
      # Filter out invalid participants
      valid_participants <- unique(allGaitParams$participant)
      current_step_outliers_data <- current_step_outliers_data[
        as.character(current_step_outliers_data$participant) %in% as.character(valid_participants), 
      ]
      
      for (i in seq_len(nrow(current_step_outliers_data))) {
        step_outlier <- current_step_outliers_data[i, ]
        
        # Find matching heel strikes for step outliers
        step_matches <- which(
          abs(allGaitParams$time - step_outlier$time) <= input$outlier_threshold &
          as.character(allGaitParams$participant) == as.character(step_outlier$participant) &
          as.numeric(allGaitParams$trialNum) == as.numeric(step_outlier$trialNum)
        )
        
        if (length(step_matches) > 0) {
          allGaitParams$outlierSteps[step_matches] <<- TRUE
        }
      }
      
      n_outlier_steps <- sum(allGaitParams$outlierSteps, na.rm = TRUE)
      showNotification(paste("Marked", n_outlier_steps, "step outliers"), 
                       type = "message", duration = 1)
    }
    
    # Update the filtered data table reference if it exists
    if (exists(".allGaitParams_dt", envir = .GlobalEnv)) {
      .allGaitParams_dt <<- NULL  # Force recreation with updated data
    }
    
    # Increment the trigger to force an update of filteredGaitParams
    refresh_trigger(refresh_trigger() + 1)
    
    # Show final summary
    n_outlier_steps <- sum(allGaitParams$outlierSteps, na.rm = TRUE)
    n_heel_bad <- sum(allGaitParams$heelStrikeBad, na.rm = TRUE)
    
    showNotification(paste("Outliers applied successfully:", n_outlier_steps, "steps marked,", 
                         n_heel_bad, "heel strikes marked for removal"), 
                   type = "message", duration = 5)
                   
    # Note: Step calculation will happen automatically during analysis
    # based on the remaining (non-removed) heel strikes
    
  }, error = function(e) {
    showNotification(paste("Error applying outliers:", e$message), type = "error", duration = 5)
    warning("Error in confirm_apply_outliers: ", e$message)
  })
})

# Handle multiple point selection (updated for new outlier system)
observeEvent(input$confirm_selection, {
  selected_data <- event_data("plotly_selected", source = "outlier_plot")
  
  if (!is.null(selected_data)) {
    current_participant <- input$participant
    current_trial <- as.numeric(input$trialNum)
    
    params <- filteredParams()
    trial_data <- params[params$participant == current_participant &
                        params$trialNum == current_trial, ]
    
    if (nrow(trial_data) == 0) {
      showNotification("No data found for current participant/trial", type = "warning")
      return()
    }
    
    if (input$outlier_mode == "heelstrike") {
      # Add multiple heel strike outliers
      new_hs_outliers <- lapply(seq_len(nrow(selected_data)), function(i) {
        clicked_time <- selected_data$x[i]
        
        closest_step <- trial_data %>%
          mutate(time_diff = abs(time - clicked_time)) %>%
          arrange(time_diff) %>%
          slice(1)
        
        if (nrow(closest_step) > 0 && closest_step$time_diff <= 0.25) {
          data.frame(
            participant = closest_step$participant,
            trialNum = closest_step$trialNum,
            time = closest_step$time,
            stringsAsFactors = FALSE
          )
        } else {
          NULL
        }
      })
      
      new_hs_outliers <- do.call(rbind, new_hs_outliers)
      
      if (!is.null(new_hs_outliers) && nrow(new_hs_outliers) > 0) {
        current_hs_outliers <- heelstrike_outliers()
        updated_outliers <- rbind(current_hs_outliers, new_hs_outliers)
        updated_outliers <- unique(updated_outliers)
        heelstrike_outliers(updated_outliers)
        showNotification(paste(nrow(new_hs_outliers), "heel strike outlier(s) added"), type = "message")
      }
    } else {
      showNotification("Multi-selection for step outliers not yet implemented", type = "message")
    }
  }
})

output$save_rds <- downloadHandler(
  filename = function() {
    file.path(dataExtraFolder, paste("allGaitParams-", Sys.Date(), ".rds", sep=""))
  },
  content = function(file) {
    saveRDS(allGaitParams, file)
  }
)
  

# Modify the reactive timer to use the user-defined interval
autoSaveTimer <- reactiveTimer(1000 * 60 * 5)  # Default to 5 minutes

# Observe changes to the autosave interval
observeEvent(input$autosave_interval, {
  invalidateLater(1000 * 60 * input$autosave_interval)
})

# Modify the autosave observer
observe({
  # Only proceed if autosave is enabled
  req(input$do_autosave)
  
  # Trigger the timer
  autoSaveTimer()
  
  isolate({
    # Get the current outliers data for both types
    hs_outliers_data <- heelstrike_outliers()
    step_outliers_data <- step_outliers()
    overwrite_mode <- input$autosave_overwrite
  })
  
  # Determine filenames and paths based on overwrite mode
  if (overwrite_mode) {
    # Overwrite default files
    hs_filename <- "outliers_heelstrikes.csv"
    step_filename <- "outliers_steps.csv"
    save_type <- "overwritten"
  } else {
    # Create timestamped files
    timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
    hs_filename <- paste0("heelstrike_outliers_autosave_", timestamp, ".csv")
    step_filename <- paste0("step_outliers_autosave_", timestamp, ".csv")
    save_type <- "saved"
  }
  
  # Define the full paths for the files
  hs_filepath <- file.path(dataExtraFolder, hs_filename)
  step_filepath <- file.path(dataExtraFolder, step_filename)
  
  # Write the CSV files
  tryCatch({
    write.csv(hs_outliers_data, hs_filepath, row.names = FALSE)
    write.csv(step_outliers_data, step_filepath, row.names = FALSE)
    
    # Show a notification with appropriate message
    if (overwrite_mode) {
      showNotification(paste("Outliers auto-saved (overwritten default files):", 
                           nrow(hs_outliers_data), "heel strikes,", 
                           nrow(step_outliers_data), "steps"), 
                     type = "message", duration = 3)
    } else {
      showNotification(paste("Outliers auto-saved (timestamped):", 
                           nrow(hs_outliers_data), "heel strikes,", 
                           nrow(step_outliers_data), "steps"), 
                     type = "message", duration = 3)
    }
  }, error = function(e) {
    showNotification(paste("Auto-save failed:", e$message), type = "error", duration = 5)
  })
  
  # Schedule the next autosave
  invalidateLater(1000 * 60 * input$autosave_interval)
})
```