#### Participant Filtering
```{r}
selectizeInput("filterParticipants", "Participants", choices = participants, selected = participants, multiple = TRUE)
div(style = "margin-top: 5px;",
  actionButton("selectAllParticipants", "Select All", class = "btn-xs btn-success"),
  actionButton("clearAllParticipants", "Clear All", class = "btn-xs btn-danger", style = "margin-left: 6px;")
)
```

```{r, context="server"}
observeEvent(input$selectAllParticipants, {
  updateSelectizeInput(session, "filterParticipants", selected = participants)
}, ignoreInit = TRUE)

observeEvent(input$clearAllParticipants, {
  updateSelectizeInput(session, "filterParticipants", selected = character(0))
}, ignoreInit = TRUE)
```

#### Trial Filtering
```{r}
defaultTrials <- c(5,10,11)
selectizeInput("filterTrials", "Trials", choices = allTrials, selected = defaultTrials, multiple = TRUE)
div(style = "margin-top: 5px;",
  actionButton("selectAllTrials", "Select All", class = "btn-xs btn-success"),
  actionButton("selectDefaultTrials", "Select Default", class = "btn-xs btn-info", style = "margin-left: 6px;"),
  actionButton("clearAllTrials", "Clear All", class = "btn-xs btn-danger", style = "margin-left: 6px;")
)
selectizeInput("filterCondition", "Condition", choices = c("control", "perturbation", "perturbation_visualization"), selected = c("control", "perturbation", "perturbation_visualization"), multiple = TRUE)
phaseChoices <- unique(unlist(default_phases))
selectizeInput("filterPhase", "Phase", choices = phaseChoices, selected = phaseChoices, multiple = TRUE)
```

```{r, context="server"}
observeEvent(input$selectAllTrials, {
  updateSelectizeInput(session, "filterTrials", selected = allTrials)
}, ignoreInit = TRUE)

observeEvent(input$clearAllTrials, {
  updateSelectizeInput(session, "filterTrials", selected = character(0))
}, ignoreInit = TRUE)

observeEvent(input$selectDefaultTrials, {
  updateSelectizeInput(session, "filterTrials", selected = c(5, 10, 11))
}, ignoreInit = TRUE)
```

#### Step Filtering
Note: slicing doesn't work completely atm. But left the old code in because I will probably need it later.
```{r}
selectizeInput("filterOutliers", "Outliers", choices = c(TRUE, FALSE), selected = c(FALSE), multiple = TRUE) # INCLUDE outliers (TRUE), INCLUDE non-outliers (FALSE)
selectizeInput("filterSuspect", "Suspect Steps", choices = c(TRUE, FALSE), selected = c(TRUE, FALSE), multiple = TRUE) # INCLUDE suspect steps (TRUE), INCLUDE non-suspect steps (FALSE)

checkboxInput("do_slicing", "Slice data", value = FALSE)
numericInput("slice_length", "Slice data (seconds)",  min = 0, max = 180, value = 9999, step = 10)
checkboxInput("remove_middle_slices", "Only keep first and last slice", value=FALSE)
```

#### Left vs Right Steps
Note: you can also average across feet here, this takes the result of both feet of 1 participant in 1 trial and averages their values to create 1 value per trial per participant for the aggregated data (this step is done last, after calculating the results of the individual feet).
```{r}
selectizeInput("filterSide", "Side", choices = c("Left", "Right"), selected = c("Left", "Right"), multiple = TRUE)
checkboxInput("avg_feet", "Average across feet", value = TRUE)
checkboxInput("add_diff", "add diffFeet columns",value = FALSE)
bsTooltip("add_diff", 
          "Calculate difference of calculated metrics (mean, sd, cv) between left and right foot and add to summary table.", 
          placement = "right", 
          trigger = "hover")
```

#### Figure Options
```{r}
numericInput("plotheight", "Plot Height (pixels)",
  min = 50, max = Inf, value = 500, step = 50)
numericInput("plotwidth", "Plot Width (pixels)",
  min = 50, max = Inf, value = 1000, step = 50)

numericInput("baseSize", "Base Size",
  min = 1, max = Inf, value = 10, step = 1)
```
