---
title: "VFD for gait variability"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll # scroll # fill
runtime: shinyrmd
---

```{r setup, include=FALSE}

# to reset:
# Kill ALL R processes (including your current session - use with caution!)
# system("pkill -f '^R$|Rterm|rsession'")
# Kill all background R processes
#system("pkill -f 'R --slave'", wait = FALSE)
#system("pkill -f 'Rscript'", wait = FALSE)
#system("pkill -f 'parallel'", wait = FALSE)

rmarkdown::shiny_prerendered_clean("workspace/index.Rmd") # we need to clear else it doesnt start the interface for whatever reason. 

# Configuration variables are now centralized in source/initialization.R

source('source/setup.R')
source('source/data_loading.R', local = FALSE)
source('source/pre_processing.R', local = FALSE)

source('source/find_foot_events.R', local = FALSE)

# =============================================================================
# FORCE REINITIALIZATION AT STARTUP
# =============================================================================
# Force a complete reinitialization of all global data and parameters
# This ensures everything is properly loaded before any calculations begin
cat("=== FORCING COMPLETE REINITIALIZATION AT STARTUP ===\n")
initialize_global_parameters()
initialize_global_data()
cat("=== REINITIALIZATION COMPLETE ===\n")
# =============================================================================

## Get Results
source('source/get_data_from_loop.R', local = FALSE)
source('source/calc_all_gait_params.R', local = FALSE)
source('source/get_questionnaire_results.R', local = FALSE)

# File paths (using centralized parameters)
agpFile <- file.path(resultsFolder, "allGaitParams.rds")
aqrFile <- file.path(resultsFolder, "allQResults.rds")
simDataFile <- file.path(resultsFolder, "allSimulationData.rds")
taskMetricsFile <- file.path(resultsFolder, "allTaskMetrics.rds")
complexityMetricsFile <- file.path(resultsFolder, "allComplexityMetrics.rds")

# Load or calculate the data using the global parallel setting
allQResults <- load_or_calculate(aqrFile, get_all_questionnaire_results)
allGaitParams <- load_or_calculate(agpFile, calc_all_gait_params, force_recalc = FALSE)

dataTypes <- setdiff(getTypes(allGaitParams), categories)

## The simulation source
source('source/profile_shapes.R', local = FALSE)
source('source/simulation_core.R', local = FALSE)
source('source/get_simulation_data.R', local = FALSE)

## Summary tables
source('source/summarize_gaitparams.R', local = FALSE)
source('source/summarize_simulation.R', local = FALSE)
source('source/complexity.R', local = FALSE)

# Load or calculate task and complexity metrics using the global parallel setting
allTaskMetrics <- load_or_calculate(taskMetricsFile, get_all_task_metrics)
allComplexityMetrics <- load_or_calculate(complexityMetricsFile, get_all_complexity_metrics)

## Plotting code
source('source/plotting.R', local = FALSE)

# for testing combining of tables
#mu_gait <- get_full_mu(allGaitParams, categories, TRUE, FALSE)
#mu_combined <- merge_mu_with_questionnaire(mu_gait, allQResults)
#mu_combined <- merge_mu_with_task(mu_gait, allTaskMetrics)
#mu_combined <- merge_mu_with_complexity(mu_combined, allComplexityMetrics)
```

Sidebar {.sidebar}
=====================================
```{r, child='pages/sidebar_dynamicDataFiltering.Rmd'}
```

```{r, context="server", local=FALSE}
source('source/get_filtered_data.R', local = TRUE)
```


```{r, context="server", local=FALSE}
session$allowReconnect(TRUE)

refresh_trigger <- reactiveVal(0) # used for refreshing filteredGaitParams after outlier update

#Custom function used to render SVG's instead of png's
renderSVG <- function(expr) {
  renderImage({
    # Dynamically get width and height from input
    dpi <- 96
    width_value <- input$plotwidth / dpi
    height_value <- input$plotheight / dpi

    file <- htmltools::capturePlot(
      expr(), tempfile(fileext = ".svg"),
      grDevices::svg,
      width = width_value, height = height_value
    )
    list(src = file, deleteFile = TRUE)
  }, deleteFile = TRUE)
}
```

Data Dictionary.md {data-orientation=columns}
=====================================
```{r, echo=FALSE}
shiny::includeMarkdown('data_dictionary.md')
```

Feet trajectories {data-orientation=columns}
=====================================
```{r, child='pages/page1_feetTrajectories.Rmd'}
```


Removed Steps
=====================================
```{r, child='pages/page2_removedSteps.Rmd'}
```


Raw tracker data {data-orientation=columns}
=====================================
```{r, child='pages/page3_rawTrackerData.Rmd'}
```


Histograms 
=====================================
```{r, child='pages/page5_histograms.Rmd'}
```

Scatterplots
=====================================
```{r, child='pages/page6_scatterplots.Rmd'}
```

Scatterplots TrialData
=====================================
```{r, child='pages/page7_trialScatterplots.Rmd'}
```

Questionnaires 
=====================================
```{r, child='pages/page8_questionnaires.Rmd'}
```

Boxplots 
=====================================
```{r, child='pages/page9_boxplots.Rmd'}
```

Statistics {data-icon="fa-chart-bar"}
=====================================
```{r, child='pages/page10_statistics.Rmd'}
```

Correlations {data-icon="fa-chart-line"}
=====================================
```{r, child='pages/page11_correlations.Rmd'}
```

Participant Summary {data-icon="fa-id-card-clip"}
=====================================
```{r, child='pages/page12_participantSummary.Rmd'}
```

Data Table {data-icon="fa-table"}
=====================================
```{r, child='pages/page13_table.Rmd'}
```

Y-Rotation Correction
=====================================
```{r, child='pages/page14_dataCorrection.Rmd'}
```

Manual Outlier Filtering
=====================================
```{r, child='pages/page16_manualOutlierFiltering.Rmd'}
```

Simulation Data {data-icon="fa-gamepad"}
=====================================
```{r, child='pages/page17_simulationData.Rmd'}
```
